---
title: "Explore raw values of WiMMed variables"
author: "AJ Perez-Luque (@ajpelu); FJ Bonet; J Herrero and R. Perez-Perez"
date: "2016 March"
output:
  md_document:
      variant: markdown_github
---

```{r metadata, echo=FALSE}
# Set working directory 

# machine <- 'ajpelu'
machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/MS/CONGRESO_EGU2016/sn_snow_egu_2016', sep='')
```

```{r packages, warning=FALSE, message=FALSE}
# Load packages 
library("raster")
library("rgdal")
library("sp")
library("plyr")
library("dplyr")
library("rasterVis") 
library("multcomp")
library("broom") # tidy output models
library("grid") # multiple plots
library("gridExtra") # multiple plots
library("ggplot2")
library("GGally")
library("pander")
library("mclust")
source("http://www.highstat.com/BGS/GAM/RCode/HighstatLibV8.R")
source(paste0(di,"/R/exportpdf.R")) # function to export raster levelplots maps as pdf
```

```{r, code=readLines(paste(di,'/R/autocaption.R', sep='')), echo=FALSE}
```


## Prepare Data

* Read snow cover indicator data and subset snow cover duration 
* Read topographic data and position (spatial) data
* Read region data (hydrological basins)

```{r}
# Read data (snow cover)
pre <- read.csv(file=paste(di, "/data/raw/PXpxXyear.csv", sep= ""), header = TRUE) 
pre_snow <- read.csv(file=paste(di, "/data/raw/PnXpxXyear.csv", sep= ""), header = TRUE) 
pre_snow_per <- read.csv(file=paste(di, "/data/raw/PorcPnXpxXyear.csv", sep= ""), header = TRUE) 
temp <- read.csv(file=paste(di, "/data/raw/TXpxXyear.csv", sep= ""), header = TRUE) 


# Join Wimmed data
join_aux1 <- inner_join(pre, pre_snow, by=c('id','year'))
join_aux2 <- inner_join(pre_snow_per, temp, by=c('id','year'))
wimmed <- inner_join(join_aux1, join_aux2, by=c('id','year'))

# Set rigth names
names(wimmed)[1] <- "nie_malla_modi_id"



# --   

# Read spatial data and Get lat/long
centroides <- rgdal::readOGR(dsn=paste(di, "/data/geoinfo", sep=""),
                             layer = "centroides_selected", verbose = FALSE)
# Select only attributes of interest and rename them
centroides <- centroides[c("id")]

# Create lat/lng by id 
xycentroides <- cbind(centroides@data, coordinates(centroides))
names(xycentroides) <- c("nie_malla_modi_id", "lon","lat")

xycentroides <- filter(xycentroides, nie_malla_modi_id %in% wimmed$nie_malla_modi_id)

# -- 

# Read Topographic data 
rawtopo <- read.csv(file=paste(di, "/data/topo_nie_malla_modis.csv", sep=""),
                    header=TRUE,
                    sep = ",") 
# function to convert radian to degree 
rad2deg <- function(rad) {(rad * 180) / (pi)} 


topo <- rawtopo %>% 
  filter(id %in% wimmed$nie_malla_modi_id) %>% 
  mutate(nie_malla_modi_id = id, 
         slope50mean_deg = rad2deg(slope50mean),
         slope50median_deg = rad2deg(slope50median),
         aspect50mean_deg = rad2deg(aspect50mean),
         aspect50median_deg = rad2deg(aspect50median)) %>%
  dplyr::select(nie_malla_modi_id, dem50mean, dem50median, slope50mean_deg, 
                slope50median_deg, aspect50mean_deg, aspect50median_deg) 
# -- 

## Hydrological basin 
basin <- read.csv(file=paste(di, "/data/derived/pixel_region.csv", sep=""),
                    header=TRUE,
                    sep = ",") 
# --


# Create objects with basis statistics for all indicators 
indicadores <- c("pre", "pre_snow", "pre_snow_per", "temp") 

# Loop to create objects with basis stats
for (i in indicadores) { 
  vnames <- c("nie_malla_modi_id", i) 
  aux <- wimmed %>%
    dplyr::select(one_of(vnames)) %>%
    mutate_(vinterest = i) %>%
    dplyr::group_by_("nie_malla_modi_id") %>%
    summarise(mean=mean(vinterest),
              sd = sd(vinterest),
              cv = raster::cv(vinterest),
              se = sd / sqrt (length(vinterest))) %>%
    inner_join(topo, by="nie_malla_modi_id") %>%
    inner_join(xycentroides, by="nie_malla_modi_id") %>%
    inner_join(basin, by="nie_malla_modi_id")
  
  assign(i, aux)
} 

```




```{r echo=FALSE}
# Set theme ggplot options
mythemeggplot <- theme_bw() + theme(panel.grid.major=element_blank(),
                       panel.grid.minor=element_blank(),
                       strip.background=element_rect(fill='white'))
```

## Explore variables with elevation 

```{r, echo=FALSE, fig.cap=fig$cap("elev_filter", "Profile of elevation"), fig.height=8 , fig.width=8}
# Pre 
plot_pre <- ggplot(pre, aes(x=dem50mean, y=mean)) +
    geom_point() +  
  mythemeggplot + 
  xlab('Elevation') + ylab('Rainfall (mm)') 

pdf(file=paste0(di, "/images/plot_elevation_pre.pdf"), height = 10, width = 10)
plot_pre
dev.off()

# Pre Snow 
plot_pre_snow <- ggplot(pre_snow, aes(x=dem50mean, y=mean)) +
    geom_point() +  mythemeggplot + 
  xlab('Elevation') + ylab('Snow Rainfall (mm)')

pdf(file=paste0(di, "/images/plot_elevation_pre_snow.pdf"), height = 10, width = 10)
plot_pre_snow
dev.off()

# Pre Snow PEr
plot_pre_snow_per <- ggplot(pre_snow_per, aes(x=dem50mean, y=mean)) +
    geom_point() +  
  mythemeggplot + 
  xlab('Elevation') + ylab('Snow Rainfall (%)') 

pdf(file=paste0(di, "/images/plot_elevation_pre_snow_per.pdf"), height = 10, width = 10)
plot_pre_snow_per
dev.off()


# Temp
plot_temp <- ggplot(temp, aes(x=dem50mean, y=mean)) +
    geom_point() +  
  mythemeggplot + 
  xlab('Elevation') + ylab('Mean Temperature') 

pdf(file=paste0(di, "/images/plot_elevation_temp.pdf"), height = 10, width = 10)
plot_temp
dev.off()


```



### Spatial pattern of the wimmed indicators 

* All these rasters are stored at `./data/derived/` 

```{r} 
# Spatial data
# Reproject to utm and m
centroides <- spTransform(centroides, CRS("+init=epsg:23030"))

# Get projection 
projection(centroides) 

# Select only attributes of interest and rename them
centroides <- centroides[c("id")]
names(centroides) <-"nie_malla_modi_id"

# Which pixels are in the snow object
pix_comunes <- match(wimmed$nie_malla_modi_id, centroides$nie_malla_modi_id)

# Create spatial objetc with centroid of Sierra Nevada
centroides_sn <- centroides[pix_comunes,]

# elevation filter
# el <- '1250'
# el_numeric <- 1250

# Loop to create raster map 
for (i in indicadores) { 
  df <- get(i)
  
  # merge MKT and spatial pixel 
  aux_spatial <- sp::merge(x=centroides_sn, y=df, by="nie_malla_modi_id")
  # aux_spatial_elev <- aux_spatial[c("nie_malla_modi_id", "dem50mean")]
  
  # raster auxiliar 
  aux_rast <- raster(aux_spatial, resolution=500)
  
  # raster of Mean values 
  mean_raster <- rasterize(aux_spatial, aux_rast, "mean", fun=mean)
  names(mean_raster) <- i # Set name of the raster layer 
  
  # raster of CV values
  cv_raster <- rasterize(aux_spatial, aux_rast, "cv", fun=mean)
  names(cv_raster) <- i 


  # assign 
  name_mean <- paste("r_mean_",i, sep="")
  name_cv <- paste("r_cv_",i, sep="")
  
  assign(name_mean, mean_raster)
  writeRaster(mean_raster, file=paste(di, "/data/derived/r_mean_", i, ".asc", sep=""), overwrite=TRUE)
  
  assign(name_cv, cv_raster)
  writeRaster(cv_raster, file=paste(di, "/data/derived/r_cv_", i, ".asc", sep=""), overwrite=TRUE)
  
}
```

## Visualization 

For each indicator we plot several maps. See `/images/raster_maps/`

### Pre

```{r, echo=FALSE, fig.cap=fig$cap("pre_mean", "Mean values of Rainfall"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
mypal <- brewer.pal(9, "Blues")

# Specify the color palette
myTheme=rasterTheme(region=mypal)


# Mean plot
lp <- levelplot(r_mean_pre, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE,
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          # at=c(-1,seq(0,225,by=25)),
          main="Rainfall (mm)")
print(lp)
exportpdf(mypdf=paste0(di, '/images/raster_maps/r_mean_pre.pdf'), lp) 


## Lat long 

# Auxiliar layer for Reproject the maps 
aux_project <- rgdal::readOGR(dsn=paste(di, "/data/geoinfo", sep=""),
                             layer = "centroides_selected", verbose = FALSE)

crs_aux_project <- projection(aux_project)

## Boundaries SN 
enp <- rgdal::readOGR(dsn=paste("/Users/", machine, "/Dropbox/carto_public/EENNPP/InfGeografica/InfVectorial/Shapes/ED50_30", sep=""),
                      layer = "EENNPP", verbose = FALSE)
# Subset limits of SN                      
sn <- subset(enp, NOMBRE == 'SIERRA NEVADA' & FIGURA == 'Espacio Natural')

# Reproject limits
sn_re <- spTransform(sn, CRS(crs_aux_project))


r_mean_pre_re <- projectRaster(r_mean_pre, crs=crs(aux_project))



lp <-levelplot(r_mean_pre_re, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE,
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          # at=c(-1,seq(0,225,by=25)),
          main="Rainfall (mm)") # + latticeExtra::layer(sp.polygons(sn_re))

print(lp)

exportpdf(mypdf=paste0(di, '/images/raster_maps/r_mean_pre_lat_long.pdf'), lp) 
```


### Pre Snow 

```{r, echo=FALSE, fig.cap=fig$cap("pre_snow_mean", "Mean values of Snowfall"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
mypal <- brewer.pal(9, "Blues")

# Specify the color palette
myTheme=rasterTheme(region=mypal)


# Mean plot
lp <- levelplot(r_mean_pre_snow, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE,
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          # at=c(-1,seq(0,225,by=25)),
          main="Snowfall (mm)")
print(lp)
exportpdf(mypdf=paste0(di, '/images/raster_maps/r_mean_pre_snow.pdf'), lp) 


## Lat long 
r_mean_pre_snow_re <- projectRaster(r_mean_pre_snow, crs=crs(aux_project))

lp <-levelplot(r_mean_pre_snow_re, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE,
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          # at=c(-1,seq(0,225,by=25)),
          main="Snowfall (mm)") # + latticeExtra::layer(sp.polygons(sn_re))

print(lp)

exportpdf(mypdf=paste0(di, '/images/raster_maps/r_mean_pre_snow_lat_long.pdf'), lp) 
```



### Pre Snow Per

```{r, echo=FALSE, fig.cap=fig$cap("pre_snow_per_mean", "Mean values of Snowfall (%)"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
mypal <- brewer.pal(9, "Blues")

# Specify the color palette
myTheme=rasterTheme(region=mypal)


# Mean plot
lp <- levelplot(r_mean_pre_snow_per, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE,
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          # at=c(seq(0,100,by=10)),
          main="Snowfall (mm)")
print(lp)
exportpdf(mypdf=paste0(di, '/images/raster_maps/r_mean_pre_snow_per.pdf'), lp) 


## Lat long 
r_mean_pre_snow_per_re <- projectRaster(r_mean_pre_snow_per, crs=crs(aux_project))

lp <-levelplot(r_mean_pre_snow_per_re, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE,
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          # at=c(-1,seq(0,225,by=25)),
          main="Snowfall (mm)") # + latticeExtra::layer(sp.polygons(sn_re))

print(lp)

exportpdf(mypdf=paste0(di, '/images/raster_maps/r_mean_pre_snow_per_lat_long.pdf'), lp) 
```


### Temp 

```{r, echo=FALSE, fig.cap=fig$cap("temp", "Mean values of Temp (ºC)"), fig.height=8 , fig.width=11}


# Mean plot
lp <- levelplot(r_mean_temp, 
          par.settings=BuRdTheme, margin=list(axis=TRUE), colorkey=TRUE,
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          # at=c(seq(0,100,by=10)),
          main="Mean Temperature (ºC)")
print(lp)
exportpdf(mypdf=paste0(di, '/images/raster_maps/r_mean_temp.pdf'), lp) 


## Lat long 
r_mean_temp_re <- projectRaster(r_mean_temp, crs=crs(aux_project))

lp <-levelplot(r_mean_temp_re, 
          par.settings=BuRdTheme, margin=list(axis=TRUE), colorkey=TRUE,
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          # at=c(-1,seq(0,225,by=25)),
          main="Mean Temperature (ºC)") # + latticeExtra::layer(sp.polygons(sn_re))

print(lp)

exportpdf(mypdf=paste0(di, '/images/raster_maps/r_mean_temp_lat_long.pdf'), lp) 
```

