---
title: "Explore raw values of snow-cover indicators"
author: "AJ Perez-Luque (@ajpelu); FJ Bonet; J Herrero and R. Perez-Perez"
date: "2016 March"
output:
  md_document:
      variant: markdown_github
---

```{r metadata, echo=FALSE}
# Set working directory 

machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/MS/CONGRESO_EGU2016/sn_snow_egu_2016', sep='')

```

```{r packages, warning=FALSE, message=FALSE}
# Load packages 
library("raster")
library("rgdal")
library("sp")
library("dplyr")
library("rasterVis") 
library("multcomp")
library("broom") # tidy output models
library("grid") # multiple plots
library("gridExtra") # multiple plots
library("ggplot2")
library("GGally")
library("pander")
source("http://www.highstat.com/BGS/GAM/RCode/HighstatLibV8.R")
```

```{r, code=readLines(paste(di,'/R/autocaption.R', sep='')), echo=FALSE}
```


## Prepare Data

* Read snow cover indicator data and subset snow cover duration 
* Read topographic data and position (spatial) data

```{r}
# Read data (snow cover)
snow <- read.csv(file=paste(di, "/data/raw/snow_sn.csv", sep= ""), header = TRUE) 
# --   

# Read spatial data and Get lat/long
centroides <- rgdal::readOGR(dsn=paste(di, "/data/geoinfo", sep=""),
                             layer = "centroides_selected", verbose = FALSE)
# Select only attributes of interest and rename them
centroides <- centroides[c("id")]

# Create lat/lng by id 
xycentroides <- cbind(centroides@data, coordinates(centroides))
names(xycentroides) <- c("nie_malla_modi_id", "lon","lat")

xycentroides <- filter(xycentroides, nie_malla_modi_id %in% snow$nie_malla_modi_id)

# -- 

# Read Topographic data 
rawtopo <- read.csv(file=paste(di, "/data/topo_nie_malla_modis.csv", sep=""),
                    header=TRUE,
                    sep = ",") 
# function to convert radian to degree 
rad2deg <- function(rad) {(rad * 180) / (pi)} 


topo <- rawtopo %>% 
  filter(id %in% snow$nie_malla_modi_id) %>% 
  mutate(nie_malla_modi_id = id, 
         slope50mean_deg = rad2deg(slope50mean),
         slope50median_deg = rad2deg(slope50median),
         aspect50mean_deg = rad2deg(aspect50mean),
         aspect50median_deg = rad2deg(aspect50median)) %>%
  dplyr::select(nie_malla_modi_id, dem50mean, dem50median, slope50mean_deg, 
                slope50median_deg, aspect50mean_deg, aspect50median_deg) 
# -- 


# Create objects with basis statistics for all indicators 
indicadores <- c("scd", "scod", "scmd", "scmc") 

# Loop to create objects with basis stats
for (i in indicadores) { 
  vnames <- c("nie_malla_modi_id", i) 
  aux <- snow %>%
    dplyr::select(one_of(vnames)) %>%
    mutate_(vinterest = i) %>%
    group_by_("nie_malla_modi_id") %>%
    summarise(mean=mean(vinterest),
              sd = sd(vinterest),
              cv = raster::cv(vinterest),
              se = sd / sqrt (length(vinterest))) %>%
    inner_join(topo, by="nie_malla_modi_id") %>%
    inner_join(xycentroides, by="nie_malla_modi_id") 
  
  assign(i, aux)
} 

```


### Spatial pattern of the snowcover indicators 

* Create raster maps of the summary stats for each indicator (`$indicator$`: `scd`, `scod`, `scmd`, `scmc`). Two raster maps will be created:
  
  * `r_mean_$indicator$`: mean values of the indicator for the pixel in the temporal serie.
* `r_cv_$indicator$`: coefficient of variation of the indicator for the pixel in the temporal serie.

* Two additional raster maps will be created, with a mask of the elevation (those pixels above 1900 *m asl*). The names of the raster are: `r_mean_$indicator$_1900` and `r_cv_$indicator$_1900`. Pixels below 1900 masl show a value of `0`. This value can be customized (change `updatevalue=-1` argument of the `mask` function). 

* All these rasters are stored at `./data/derived/` 

```{r} 
# Spatial data
# Reproject to utm and m
centroides <- spTransform(centroides, CRS("+init=epsg:23030"))

# Get projection 
projection(centroides) 

# Select only attributes of interest and rename them
centroides <- centroides[c("id")]
names(centroides) <-"nie_malla_modi_id"

# Which pixels are in the snow object
pix_comunes <- match(snow$nie_malla_modi_id, centroides$nie_malla_modi_id)

# Create spatial objetc with centroid of Sierra Nevada
centroides_sn <- centroides[pix_comunes,]


# Loop to create raster map 
for (i in indicadores) { 
  df <- get(i)
  
  # merge MKT and spatial pixel 
  aux_spatial <- sp::merge(x=centroides_sn, y=df, by="nie_malla_modi_id")
  aux_spatial_elev <- aux_spatial[c("nie_malla_modi_id", "dem50mean")]
  
  # raster auxiliar 
  aux_rast <- raster(aux_spatial, resolution=500)
  
  # raster of Mean values 
  mean_raster <- rasterize(aux_spatial, aux_rast, "mean", fun=mean)
  names(mean_raster) <- i # Set name of the raster layer 
  
  # raster of CV values
  cv_raster <- rasterize(aux_spatial, aux_rast, "cv", fun=mean)
  names(cv_raster) <- i 
  
  # raster of Elevation
  elev_raster <- rasterize(aux_spatial_elev, aux_rast, "dem50mean", fun=mean)
  elev_raster1900 <- elev_raster
  elev_raster1900[elev_raster1900 < 1900] <- NA
  
  # Mask by 1900
  mean_raster1900 <- mask(mean_raster, elev_raster1900, updatevalue=0)
  cv_raster1900 <- mask(cv_raster, elev_raster1900, updatevalue=0)
  
  
  # assign 
  name_mean <- paste("r_mean_",i, sep="")
  name_cv <- paste("r_cv_",i, sep="")
  name_mean1900 <- paste("r_mean_",i,"_1900", sep="")
  name_cv1900 <- paste("r_cv_",i,"_1900", sep="")
  
  assign(name_mean, mean_raster)
  writeRaster(mean_raster, file=paste(di, "/data/derived/r_mean_", i, ".asc", sep=""), overwrite=TRUE)
  
  assign(name_cv, cv_raster)
  writeRaster(cv_raster, file=paste(di, "/data/derived/r_cv_", i, ".asc", sep=""), overwrite=TRUE)
  
  assign(name_mean1900, mean_raster1900)
  writeRaster(mean_raster1900, file=paste(di, "/data/derived/r_mean_", i, "_1900", ".asc", sep=""), overwrite=TRUE)
  
  assign(name_cv1900, cv_raster1900)
  writeRaster(cv_raster1900, file=paste(di, "/data/derived/r_cv_", i, "_1900", ".asc", sep=""), overwrite=TRUE)
}
```

## Visualization of the Snow Cover indicators 

### Snow Cover Duration

```{r, echo=FALSE, fig.cap=fig$cap("scd_mean", "Mean values of Snow cover duration"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
# mypal <- brewer.pal(9, "Blues")
# Add a colour for 0 value
# mypal <- c("#D2D2D2", mypal)

mypal <- c("#D2D2D2", "#C6DBEF", "#9ECAE1", "#6BAED6", "#4292C6", "#2171B5", "#08519C", "#08306B")

# Specify the color palette
myTheme=rasterTheme(region=mypal)
# myTheme$panel.background$col <- 'gray'

# Mean plot
levelplot(r_mean_scd_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=c(0,seq(50,250,by=25)),
          main="Snow Cover Duration (mean values, days)")

```

```{r, echo=FALSE, fig.cap=fig$cap("scd_cv", "Coefficient of Variation of values of Snow cover duration"), fig.height=8 , fig.width=11}
# CV plot
mypal_cv <- brewer.pal(9, "Oranges")
mypal_cv <- c("#D2D2D2", mypal_cv)

myTheme_cv=rasterTheme(region=mypal_cv)

levelplot(r_cv_scd_1900, 
          par.settings=myTheme_cv, margin=list(axis=TRUE), colorkey=TRUE, 
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=seq(0,100,by=10),
          main="Snow Cover Duration (CV)")
```


### Snow Cover Onset Date

```{r, echo=FALSE, fig.cap=fig$cap("scod_mean", "Mean values of Snow cover Onset Date"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
mypal <- brewer.pal(9, "YlOrRd")
# Add a colour for 0 value
mypal <- c("#D2D2D2", mypal)
# Specify the color palette
myTheme=rasterTheme(region=mypal)


levelplot(r_mean_scod_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          #cut=20,
          at=c(0,seq(15,105,by=7)),
          main="Snow Cover Onset Date (mean values, hydrological day)") 

levelplot(r_mean_scod_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          main="Snow Cover Onset Date (mean values, hydrological day)") 
```

```{r, echo=FALSE, fig.cap=fig$cap("scod_cv", "Coefficient of Variation of values of Snow cover Onset date"), fig.height=8 , fig.width=11}
# CV plot
mypal_cv <- brewer.pal(9, "Oranges")
mypal_cv <- c("#D2D2D2", mypal_cv)

myTheme_cv=rasterTheme(region=mypal_cv)

levelplot(r_cv_scod_1900, 
          par.settings=myTheme_cv, margin=list(axis=TRUE), colorkey=TRUE, 
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=seq(0,100,by=10),
          main="Snow Cover Onset Date (CV)")
```

### Snow Cover Melting Date

```{r, echo=FALSE, fig.cap=fig$cap("scmd_mean", "Mean values of Snow cover Melting Date"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
mypal <- rev(brewer.pal(9, "Greens"))
# Add a colour for 0 value
mypal <- c("#D2D2D2", mypal)
# Specify the color palette
myTheme=rasterTheme(region=mypal)


levelplot(r_mean_scmd_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          #cut=20,
          at=c(0,seq(140,280,by=20)),
          main="Snow Cover Melting Date (mean values, hydrological day)") 

levelplot(r_mean_scmd_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          main="Snow Cover Melting Date (mean values, hydrological day)", 
          at=c(0,seq(140,280,by=5)))
```


```{r, echo=FALSE, fig.cap=fig$cap("scmd_cv", "Coefficient of Variation of values of Snow cover Melting date"), fig.height=8 , fig.width=11}
# CV plot
mypal_cv <- brewer.pal(9, "Oranges")
mypal_cv <- c("#D2D2D2", mypal_cv)

myTheme_cv=rasterTheme(region=mypal_cv)

levelplot(r_cv_scmd_1900, 
          par.settings=myTheme_cv, margin=list(axis=TRUE), colorkey=TRUE, 
          #contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          #at=seq(0,50,by=7.5),
          main="Snow Cover Melting Date (CV)")
```

### Snow Cover Melting Cycles

```{r, echo=FALSE, fig.cap=fig$cap("scmc_mean", "Mean values of Snow cover Melting Date"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
mypal <- brewer.pal(5, "Purples")
# Add a colour for 0 value
mypal <- c("#D2D2D2", mypal)
# Specify the color palette
myTheme=rasterTheme(region=mypal)


levelplot(r_mean_scmc_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=seq(0,5,by=1),
          main="Snow Cover Melting cycles (mean values, number)") 

levelplot(r_mean_scmc_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          main="Snow Cover Melting cycles (mean values, number)")
```

```{r, echo=FALSE, fig.cap=fig$cap("scmc_cv", "Coefficient of Variation of values of Snow cover Melting cycles"), fig.height=8 , fig.width=11}
# CV plot
mypal_cv <- brewer.pal(9, "Oranges")
mypal_cv <- c("#D2D2D2", mypal_cv)

myTheme_cv=rasterTheme(region=mypal_cv)

levelplot(r_cv_scmc_1900, 
          par.settings=myTheme_cv, margin=list(axis=TRUE), colorkey=TRUE, 
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=seq(0,100,by=10),
          main="Snow Cover Melting Date (CV)")
```

# Explore patterns of raw values

## Snow Cover Onset Dates 

* Analyze the effect of aspect, slope and longtitude on snow cover onset date

```{r, echo=FALSE}
# Subset data and classify elevation and aspect
scod1900 <- scod %>% filter(dem50mean > 1900) %>%
  mutate(dem50mean_group = cut(dem50mean, 
                         breaks = seq(from=0, to=3500, by=250),
                         labels = c("0-250", "251-500","501-750","751-1000",
                                    "1001-1250","1251-1500","1501-1750","1751-2000",
                                    "2001-2250","2251-2500","2501-2750", "2751-3000",
                                    "3001-3250", "3251-3500")),
         aspect50mean_deg_group = cut(aspect50mean_deg, 
                                      breaks= c(22.5, 67.5, 112.5, 157.5, 
                                                202.5, 247.5, 292.5, 337.5, 361),
                                      labels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW"))) 


### Model 
m_scod <- glm(mean ~ lon + aspect50mean_deg_group + slope50mean_deg, data=scod1900)

# Remove na level of aspect 
df <- scod1900 %>% na.omit() 

panderOptions('round', 4)
pander(format(tidy(drop1(m_scod, test="Chi")), digits=4))
```

* What about the Variance Inflation factors? 


```{r, echo=FALSE}
# Exploring relationship
pander(as.data.frame(corvif(df[,c('lon','aspect50mean_deg_group','slope50mean_deg')])))
detach(package:pander, unload=TRUE)
```



```{r, message=FALSE, echo=FALSE, fig.cap=fig$cap("ggpair_scod", "Explore relationships of covariates (Snow cover onset dates"), fig.height=12 , fig.width=11}
# See vignettes at GGally package
# See also http://stackoverflow.com/questions/30858337/how-to-customize-lines-in-ggpairs-ggally


# Define a customized lower part
lowerFn <- function(data, mapping, method = "lm", ...) {
  p <- ggplot2::ggplot(data = data, mapping = mapping) +
    ggplot2::geom_point(colour = "gray") +
    ggplot2::geom_smooth(method = method , color = "red", ...)
  p
}

# Plot 
ggpairs(df, 
        columns = c("lon","aspect50mean_deg_group","slope50mean_deg"),
        columnLabels = c('Longitude', 'Aspect', 'Slope'),
        lower = list(continuous = wrap(lowerFn, method="lm"),
                     combo = wrap("facethist", fill="blue")),
        upper = list(continuous = wrap("cor", size=11), combo='box'),
        diag = list(discrete = wrap("barDiag", fill= 'blue'))) +
   theme_bw() 
```        




```{r, echo=FALSE, fig.cap=fig$cap("scod_tukey_aspect", "Snow cover onset dates by aspect"), fig.height=5 , fig.width=7, warning=FALSE, message=FALSE}
# Plot of Snow Cover Onset dates by aspect 

# Compute multiple comparisons 
tuk <- glht(m_scod, linfct = mcp(aspect50mean_deg_group= "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(aspect50mean_deg_group = as.factor(lhs),
         tukey_aspect = letters) 

# Specify theme
mythemeggplot <- theme(panel.grid.major=element_blank(),
                       panel.grid.minor=element_blank(),
                       strip.background=element_rect(fill='white'))
  
# Plot 
ggplot(df, aes(x=aspect50mean_deg_group, y=mean)) +geom_boxplot() + 
  geom_text(data=df_letter, aes(label = tukey_aspect, x=aspect50mean_deg_group, y=100),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot + 
  ylab('Snow cover onset date (hydrological day)') +
  xlab('aspect')  
```




```{r echo=FALSE, fig.cap=fig$cap("scod_tukey_aspect_effects", "Snow cover onset dates by aspect (effect sizes"), fig.height=8 , fig.width=11}
# Effects comparison (aspect)
CI <- confint(tuk)

ggplot(mapping = aes(lhs, estimate)) +
  geom_linerange(aes(ymin = lwr, ymax = upr), data = CI) +
  geom_point(aes(size = p), data = summary(tuk)) +
  scale_size(trans = "reverse") + theme_bw() + 
  coord_flip()
```



```{r echo=FALSE, fig.cap=fig$cap("scod_lon_slope", "Snow cover onset dates by Longitude and Slope"), fig.height=5 , fig.width=8}
y_lab <- ylab('Snow cover onset date (hydrological day)') 


x_lab <- xlab('Longitude')
a <- ggplot(scod1900, aes(x=lon, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method='lm') +
  theme_bw() + mythemeggplot + x_lab + y_lab
  
x_lab <- xlab('Slope (degrees)') 
b <- ggplot(scod1900, aes(x=slope50mean_deg, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method='lm') +
    theme_bw() + mythemeggplot + x_lab + y_lab

grid.arrange(a, b, ncol=2)
```

## Snow Cover Melting Dates 

```{r, echo=FALSE}
# Subset data and classify elevation and aspect
scmd1900 <- scmd %>% filter(dem50mean > 1900) %>%
  mutate(dem50mean_group = cut(dem50mean, 
                         breaks = seq(from=0, to=3500, by=250),
                         labels = c("0-250", "251-500","501-750","751-1000",
                                    "1001-1250","1251-1500","1501-1750","1751-2000",
                                    "2001-2250","2251-2500","2501-2750", "2751-3000",
                                    "3001-3250", "3251-3500")),
         aspect50mean_deg_group = cut(aspect50mean_deg, 
                                      breaks= c(22.5, 67.5, 112.5, 157.5, 
                                                202.5, 247.5, 292.5, 337.5, 361),
                                      labels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW"))) 


### Model 
m_scmd <- glm(mean ~ dem50mean + lon + aspect50mean_deg_group + slope50mean_deg, data=scmd1900)

# Remove na level of aspect 
df <- scmd1900 %>% na.omit() 

library("pander")
pander(format(tidy(drop1(m_scmd, test="Chi")), digits=4))
```


* What about the Variance Inflation factors? 


```{r, echo=FALSE}
# Exploring relationship
pander(as.data.frame(corvif(df[,c('dem50mean','lon','aspect50mean_deg_group','slope50mean_deg')])))
detach(package:pander, unload=TRUE)
```



```{r, message=FALSE, echo=FALSE, fig.cap=fig$cap("ggpair_scmd", "Explore relationships of covariates (Snow cover melting dates"), fig.height=12 , fig.width=11}
# Plot 
ggpairs(df, 
        columns = c("dem50mean","lon","aspect50mean_deg_group","slope50mean_deg"),
        columnLabels = c('Elevation','Longitude', 'Aspect', 'Slope'),
        lower = list(continuous = wrap(lowerFn, method="lm"),
                     combo = wrap("facethist", fill="blue")),
        upper = list(continuous = wrap("cor", size=11), combo='box'),
        diag = list(discrete = wrap("barDiag", fill= 'blue'))) +
   theme_bw() 
```        



```{r, echo=FALSE, fig.cap=fig$cap("scmd_tukey_aspect", "Snow cover melting dates by aspect"), fig.height=5 , fig.width=7, warning=FALSE, message=FALSE}
# Plot of Snow Cover Melting dates by aspect 

# Compute multiple comparisons 
tuk <- glht(m_scmd, linfct = mcp(aspect50mean_deg_group= "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(aspect50mean_deg_group = as.factor(lhs),
         tukey_aspect = letters) 

# Plot 
ggplot(df, aes(x=aspect50mean_deg_group, y=mean)) +geom_boxplot() + 
  geom_text(data=df_letter, aes(label = tukey_aspect, x=aspect50mean_deg_group, y=135),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot +
  ylab('Snow cover onset date (hydrological day)') +
  xlab('aspect') 
```




```{r echo=FALSE, fig.cap=fig$cap("scmd_tukey_aspect_effects", "Snow cover melting dates by aspect (effect sizes"), fig.height=8 , fig.width=11}
# Effects comparison (aspect)
CI <- confint(tuk)

ggplot(mapping = aes(lhs, estimate)) +
  geom_linerange(aes(ymin = lwr, ymax = upr), data = CI) +
  geom_point(aes(size = p), data = summary(tuk)) +
  scale_size(trans = "reverse") + theme_bw() + 
  coord_flip()
```




```{r echo=FALSE, fig.cap=fig$cap("scmd_lon_slope_elev", "Snow cover melting dates by Longitude, Slope and Elevation"), fig.height=6, fig.width=11}
x_lab <- xlab('Longitude')
y_lab <- ylab('Snow cover melting date (hydrological day)')

a <- ggplot(scmd1900, aes(x=lon, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method="lm", col="gray") +
  geom_smooth(method='loess') +
  theme_bw() + mythemeggplot + x_lab + y_lab
  

x_lab <- xlab('Slope (degrees)')  
b <- ggplot(scmd1900, aes(x=slope50mean_deg, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method="lm") +
  theme_bw() + mythemeggplot + x_lab + y_lab
  
x_lab <- xlab('Elevation (m)')  
c <- ggplot(scmd1900, aes(x=dem50mean, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method="lm") +
  theme_bw() + mythemeggplot + x_lab + y_lab 
  
grid.arrange(a, b, c, ncol=3)
```


## Snow Cover Melting Cycles

```{r, echo=FALSE}
# Subset data and classify elevation and aspect
scmc1900 <- scmc %>% filter(dem50mean > 1900) %>%
  mutate(dem50mean_group = cut(dem50mean, 
                         breaks = seq(from=0, to=3500, by=250),
                         labels = c("0-250", "251-500","501-750","751-1000",
                                    "1001-1250","1251-1500","1501-1750","1751-2000",
                                    "2001-2250","2251-2500","2501-2750", "2751-3000",
                                    "3001-3250", "3251-3500")),
         aspect50mean_deg_group = cut(aspect50mean_deg, 
                                      breaks= c(22.5, 67.5, 112.5, 157.5, 
                                                202.5, 247.5, 292.5, 337.5, 361),
                                      labels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW"))) 


### Model 
m_scmc <- glm(mean ~ dem50mean + lon + aspect50mean_deg_group + slope50mean_deg, data=scmc1900)

# Remove na level of aspect 
df <- scmc1900 %>% na.omit() 

library("pander")
pander(format(tidy(drop1(m_scmc, test="Chi")), digits=4))
```


* What about the Variance Inflation factors? 


```{r, echo=FALSE}
# Exploring relationship
pander(as.data.frame(corvif(df[,c('dem50mean','lon','aspect50mean_deg_group','slope50mean_deg')])))
detach(package:pander, unload=TRUE)
```



```{r, message=FALSE, echo=FALSE, fig.cap=fig$cap("ggpair_scmc", "Explore relationships of covariates (Snow cover melting dates"), fig.height=12 , fig.width=11}
# Plot 
ggpairs(df, 
        columns = c("dem50mean","lon","aspect50mean_deg_group","slope50mean_deg"),
        columnLabels = c('Elevation','Longitude', 'Aspect', 'Slope'),
        lower = list(continuous = wrap(lowerFn, method="lm"),
                     combo = wrap("facethist", fill="blue")),
        upper = list(continuous = wrap("cor", size=11), combo='box'),
        diag = list(discrete = wrap("barDiag", fill= 'blue'))) +
   theme_bw() 
```        



```{r, echo=FALSE, fig.cap=fig$cap("scmc_tukey_aspect", "Snow cover melting dates by aspect"), fig.height=5 , fig.width=7, warning=FALSE, message=FALSE}
# Plot of Snow Cover Melting dates by aspect 

# Compute multiple comparisons 
tuk <- glht(m_scmc, linfct = mcp(aspect50mean_deg_group= "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(aspect50mean_deg_group = as.factor(lhs),
         tukey_aspect = letters) 

# Plot 
ggplot(df, aes(x=aspect50mean_deg_group, y=mean)) +geom_boxplot() + 
  geom_text(data=df_letter, aes(label = tukey_aspect, x=aspect50mean_deg_group, y=5),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot +
  ylab('Snow cover melting cycles (n)') +
  xlab('aspect') 
```




```{r echo=FALSE, fig.cap=fig$cap("scmc_tukey_aspect_effects", "Snow cover melting dates by aspect (effect sizes"), fig.height=8 , fig.width=11}
# Effects comparison (aspect)
CI <- confint(tuk)

ggplot(mapping = aes(lhs, estimate)) +
  geom_linerange(aes(ymin = lwr, ymax = upr), data = CI) +
  geom_point(aes(size = p), data = summary(tuk)) +
  scale_size(trans = "reverse") + theme_bw() + 
  coord_flip()
```




```{r echo=FALSE, fig.cap=fig$cap("scmc_lon_slope_elev", "Snow cover melting dates by Longitude, Slope and Elevation"), fig.height=6, fig.width=11}
x_lab <- xlab('Longitude')
y_lab <- ylab('Snow cover melting cycles (n)')

a <- ggplot(scmc1900, aes(x=lon, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method='lm') +
  theme_bw() + mythemeggplot + x_lab + y_lab
  

x_lab <- xlab('Slope (degrees)')  
b <- ggplot(scmc1900, aes(x=slope50mean_deg, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method="lm") +
  theme_bw() + mythemeggplot + x_lab + y_lab
  
x_lab <- xlab('Elevation (m)')  
c <- ggplot(scmc1900, aes(x=dem50mean, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method="lm", col="gray") +
  geom_smooth(method="loess") +
  theme_bw() + mythemeggplot + x_lab + y_lab 
  
grid.arrange(a, b, c, ncol=3)
```



:red_circle: por aqui vas carpinto

```{r, echo=FALSE, eval=FALSE}
ggplot(df, aes(x=aspect, y=mean)) + 
  geom_bar(position=position_dodge(), stat="identity", fill='#084594') +
  geom_errorbar(aes(ymax = mean + sd, ymin=mean - sd),
                  width=.25, col= '#084594', 
                  position=position_dodge(.9)) + 
  theme_bw() + 
  ylab('Snow cover onset date (hydrological day)') +
  xlab('aspect') + 
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.background=element_rect(fill='white')) 
```





