---
title: "Explore Trends of WiMMed Indicators"
author: "AJ Perez-Luque (@ajpelu); FJ Bonet; J Herrero and R. Perez-Perez"
date: "2016 March"
output: 
    md_document:
      variant: markdown_github
      
---
## Prepare Data

1. Read data of Mann-Kendal Sen-Slope for each pixels and WiMMed indicators 
2. Read data of topographic variable: 
    + Convert radian to deg. 
    + Create categorical variable for elevation (250 m)
    + Classify aspect into 8 categories 
3. Read data from hydrological basin 
4. Read spatial data: 
    + Select only centroides of interest 
5. Create two dataframes: 
    + **Full Dataframe** with all variables and all pixels (`fulldf`)
    + :red_circle: In this exploration we don't filter the data by elevation 
    


```{r metadata, echo=FALSE}
# Set working directory 

# machine <- 'ajpelu'
machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/MS/CONGRESO_EGU2016/sn_snow_egu_2016', sep='')
```


```{r packages, warning=FALSE, echo=FALSE, message=FALSE}
# Load packages 
library("raster")
library("rgdal")
library("sp")
library("plyr")
library("dplyr")
library("ggplot2")
library("reshape2")
library("multcomp")
library("pander")
library("broom")
library("mgcv")
library("grid") # multiple plots
library("gridExtra") # multiple plots
source('https://raw.githubusercontent.com/ajpelu/ajpeluR/master/R/exploreMKTS.R')
```

```{r data}
# Trend analysis data
# Define name of indicators (see variables names)
# Use annual and season aggregation 
indicadores <- c("pre", "pre_snow", "pre_snow_per", "temp", 
                 "pnau", "pnsp", "pnsu", "pnwi",
                 "preau", "presp", "presu", "prewi",
                 "tempau", "tempsp", "tempsu", "tempwi")

# Loop to read files 
for (j in indicadores){ 
  aux <- read.csv(file=paste(di, "/data/derived/", j, ".csv", sep= ""),
              header = TRUE,
              sep = ',')
  assign(j, aux)
}
# --

# Define pixels of interes 
pixels_interes <- pre$nie_malla_modi_id

# Read Topographic data 
rawtopo <- read.csv(file=paste(di, "/data/topo_nie_malla_modis.csv", sep=""),
                    header=TRUE,
                    sep = ",") 

# function to convert radian to degree 
rad2deg <- function(rad) {(rad * 180) / (pi)} 


topo <- rawtopo %>% 
  filter(id %in% pixels_interes) %>% 
  mutate(nie_malla_modi_id = id, 
         slope50mean_deg = rad2deg(slope50mean),
         slope50median_deg = rad2deg(slope50median),
         aspect50mean_deg = rad2deg(aspect50mean),
         aspect50median_deg = rad2deg(aspect50median)) %>%
  dplyr::select(nie_malla_modi_id, dem50mean, dem50median, slope50mean_deg, 
                slope50median_deg, aspect50mean_deg, aspect50median_deg) 

## Create interval variables (250 m) for dem; and classify aspect into 8 categories 
topo <- topo %>% 
  mutate(dem50mean_group = cut(dem50mean, 
                         breaks = seq(from=0, to=3500, by=250),
                         labels = c("0-250", "251-500","501-750","751-1000",
                                    "1001-1250","1251-1500","1501-1750","1751-2000",
                                    "2001-2250","2251-2500","2501-2750", "2751-3000",
                                    "3001-3250", "3251-3500")),
         aspect50mean_deg_group = cut(aspect50mean_deg, 
                                      breaks= c(22.5, 67.5, 112.5, 157.5, 202.5, 247.5, 292.5, 337.5, 359.5),
                                      labels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW")))


# --

# Read spatial data and Get lat/long
centroides <- rgdal::readOGR(dsn=paste(di, "/data/geoinfo", sep=""),
                             layer = "centroides_selected", verbose = FALSE)
# Select only attributes of interest and rename them
centroides <- centroides[c("id")]

# Create lat/lng by id 
xycentroides <- cbind(centroides@data, coordinates(centroides))
names(xycentroides) <- c("nie_malla_modi_id", "lon","lat")

xycentroides <- filter(xycentroides, nie_malla_modi_id %in% pixels_interes)
# -- 

# Hydrological basin 
basin <- read.csv(file=paste(di, "/data/derived/pixel_region.csv", sep=""),
                    header=TRUE,
                    sep = ",") 
# --


# Create un dataframe con todos los datos
fulldf <- topo %>% 
  inner_join(pre,  by=c("nie_malla_modi_id")) %>% 
  inner_join(pre_snow,  by=c("nie_malla_modi_id")) %>% 
  inner_join(pre_snow_per,  by=c("nie_malla_modi_id")) %>% 
  inner_join(temp,  by=c("nie_malla_modi_id")) %>% 
  inner_join(pnau,  by=c("nie_malla_modi_id")) %>% 
  inner_join(pnsp,  by=c("nie_malla_modi_id")) %>% 
  inner_join(pnsu,  by=c("nie_malla_modi_id")) %>% 
  inner_join(pnwi,  by=c("nie_malla_modi_id")) %>% 
  inner_join(preau,  by=c("nie_malla_modi_id")) %>% 
  inner_join(presp,  by=c("nie_malla_modi_id")) %>% 
  inner_join(presu,  by=c("nie_malla_modi_id")) %>% 
  inner_join(prewi,  by=c("nie_malla_modi_id")) %>% 
  inner_join(tempau,  by=c("nie_malla_modi_id")) %>% 
  inner_join(tempsp,  by=c("nie_malla_modi_id")) %>% 
  inner_join(tempsu,  by=c("nie_malla_modi_id")) %>% 
  inner_join(tempwi,  by=c("nie_malla_modi_id")) %>% 
  inner_join(xycentroides, by="nie_malla_modi_id") %>%
  inner_join(basin, by="nie_malla_modi_id")


# Create subset of pixels above 1250 
fulldf1250 <- fulldf %>% 
  filter(dem50mean > 1250)  
```


# Explore general pattern of trends  
```{r}
exp_pre <- exploreMKTS(pre, alpha=0.05)
exp_pre_snow <- exploreMKTS(pre_snow, alpha=0.05)
exp_pre_snow_per <- exploreMKTS(pre_snow_per, alpha=0.05)
exp_temp <- exploreMKTS(temp, alpha=0.05)


# Create table with all resuls 
# tau_pos
tau_pos <- rbind(
  cbind(exp_pre$summary[1,c(2,3)],variable='pre'),
  cbind(exp_pre_snow$summary[1,c(2,3)],variable='pre_snow'),
  cbind(exp_pre_snow_per$summary[1,c(2,3)],variable='pre_snow_per'),
  cbind(exp_temp$summary[1,c(2,3)],variable='temp'))
  names(tau_pos) <- c('tau_pos_n','tau_pos_%','variable')

# tau_pos_sig
tau_pos_sig <- rbind(
  cbind(exp_pre$summary[2,c(2,3)],variable='pre'),
  cbind(exp_pre_snow$summary[2,c(2,3)],variable='pre_snow'),
  cbind(exp_pre_snow_per$summary[2,c(2,3)],variable='pre_snow_per'),
  cbind(exp_temp$summary[2,c(2,3)],variable='temp'))
names(tau_pos_sig) <- c('tau_pos_sig_n','tau_pos_sig_%','variable')


# tau_neg
tau_neg <- rbind(
  cbind(exp_pre$summary[3,c(2,3)],variable='pre'),
  cbind(exp_pre_snow$summary[3,c(2,3)],variable='pre_snow'),
  cbind(exp_pre_snow_per$summary[3,c(2,3)],variable='pre_snow_per'),
  cbind(exp_temp$summary[3,c(2,3)],variable='temp'))
names(tau_neg) <- c('tau_neg_n','tau_neg_%','variable')

# tau_neg_sig
tau_neg_sig <- rbind(
  cbind(exp_pre$summary[4,c(2,3)],variable='pre'),
  cbind(exp_pre_snow$summary[4,c(2,3)],variable='pre_snow'),
  cbind(exp_pre_snow_per$summary[4,c(2,3)],variable='pre_snow_per'),
  cbind(exp_temp$summary[4,c(2,3)],variable='temp'))
names(tau_neg_sig) <- c('tau_neg_sig_n','tau_neg_sig_%','variable')

# Table with all the results 
summ_explore <- join(join(tau_pos, tau_pos_sig, by='variable'),
     join(tau_neg, tau_neg_sig, by='variable'), by='variable')
# Reorder variables
summ_explore <- summ_explore[,c(3,1:2,4:9)]
summ_explore
pander(summ_explore, caption= 'Summary trends all Sierra Nevada')
```

# Explore trends (>1250)

```{r} 
# Get nie_malla_modi_id + elev
topo_elev <- topo %>% 
  select(nie_malla_modi_id, dem50mean)

pre1250 <- pre %>% 
  inner_join(topo_elev, by = 'nie_malla_modi_id') %>% 
  filter(dem50mean > 1250)

pre_snow1250 <- pre_snow %>% 
  inner_join(topo_elev, by = 'nie_malla_modi_id') %>% 
  filter(dem50mean > 1250)

pre_snow_per1250 <- pre_snow_per %>% 
  inner_join(topo_elev, by = 'nie_malla_modi_id') %>% 
  filter(dem50mean > 1250)

temp1250 <- temp %>% 
  inner_join(topo_elev, by = 'nie_malla_modi_id') %>% 
  filter(dem50mean > 1250)


exp_pre1250 <- exploreMKTS(pre1250, alpha=0.05)
exp_pre_snow1250 <- exploreMKTS(pre_snow1250, alpha=0.05)
exp_pre_snow_per1250 <- exploreMKTS(pre_snow_per1250, alpha=0.05)
exp_temp1250 <- exploreMKTS(temp1250, alpha=0.05)


# Create table with all resuls 
# tau_pos
tau_pos <- rbind(
  cbind(exp_pre1250$summary[1,c(2,3)],variable='pre'),
  cbind(exp_pre_snow1250$summary[1,c(2,3)],variable='pre_snow'),
  cbind(exp_pre_snow_per1250$summary[1,c(2,3)],variable='pre_snow_per'),
  cbind(exp_temp1250$summary[1,c(2,3)],variable='temp'))
names(tau_pos) <- c('tau_pos_n','tau_pos_%','variable')

# tau_pos_sig
tau_pos_sig <- rbind(
  cbind(exp_pre1250$summary[2,c(2,3)],variable='pre'),
  cbind(exp_pre_snow1250$summary[2,c(2,3)],variable='pre_snow'),
  cbind(exp_pre_snow_per1250$summary[2,c(2,3)],variable='pre_snow_per'),
  cbind(exp_temp1250$summary[2,c(2,3)],variable='temp'))
names(tau_pos_sig) <- c('tau_pos_sig_n','tau_pos_sig_%','variable')


# tau_neg
tau_neg <- rbind(
  cbind(exp_pre1250$summary[3,c(2,3)],variable='pre'),
  cbind(exp_pre_snow1250$summary[3,c(2,3)],variable='pre_snow'),
  cbind(exp_pre_snow_per1250$summary[3,c(2,3)],variable='pre_snow_per'),
  cbind(exp_temp1250$summary[3,c(2,3)],variable='temp'))
names(tau_neg) <- c('tau_neg_n','tau_neg_%','variable')

# tau_neg_sig
tau_neg_sig <- rbind(
  cbind(exp_pre1250$summary[4,c(2,3)],variable='pre'),
  cbind(exp_pre_snow1250$summary[4,c(2,3)],variable='pre_snow'),
  cbind(exp_pre_snow_per1250$summary[4,c(2,3)],variable='pre_snow_per'),
  cbind(exp_temp1250$summary[4,c(2,3)],variable='temp'))
names(tau_neg_sig) <- c('tau_neg_sig_n','tau_neg_sig_%','variable')

# Table with all the results 
summ_explore <- join(join(tau_pos, tau_pos_sig, by='variable'),
                     join(tau_neg, tau_neg_sig, by='variable'), by='variable')
# Reorder variables
summ_explore <- summ_explore[,c(3,1:2,4:9)]
summ_explore
pander(summ_explore, caption= 'Summary trends Sierra Nevada (>1250)')
```



# Explore trends of annual variables by basin

We explore the pattern of snow-cover indicators trends by hydrological basin. 

```{r, echo=FALSE}
# Set theme ggplot options
mythemeggplot <- theme_bw() + theme(panel.grid.major=element_blank(),
                       panel.grid.minor=element_blank(),
                       strip.background=element_rect(fill='white'))
```



```{r}
## Summary statistics 
misvariables<- c('tau_pre', 'sen_slope_pre', 'tau_pre_snow', 'sen_slope_pre_snow',
                 'tau_pre_snow_per', 'sen_slope_pre_snow_per', 'tau_temp', 'sen_slope_temp')

variable_agrupa <- 'basin_name' 
prefijo  <- 'basin_' 
df <- fulldf

stats_basin <- data.frame() 
  
for (i in misvariables){
  vnames <- c(variable_agrupa, i)
  aux <- df %>%
    dplyr::select(one_of(vnames)) %>%
    mutate_(vinterest = i) %>%
    group_by_(.dots=variable_agrupa) %>%
    summarise(mean=mean(vinterest),
              sd = sd(vinterest),
              se = sd / sqrt (length(vinterest)))
  aux <- aux %>% 
    mutate(variable = i)
  
  stats_basin <- rbind(stats_basin, aux)
  # assign(paste0(prefijo,i), aux)
   
}
```


### Pre 

```{r}
df_letter_aux <- data.frame() 
df <- fulldf
```

```{r}
# Tau pre 
# ANOVA 
variable <- 'tau_pre'
my_ylab <- 'Tau Pre'


myformula <- formula(paste(variable, '~ basin_name', sep=''))
mod <- aov(myformula, data=df)
pander(tidy(mod), caption=my_ylab)

## Multiple comparisons 
tuk <- glht(mod, linfct = mcp(basin_name = "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(basin_name = as.factor(lhs),
         tukey_basin_name = letters) %>%
  mutate(variable = variable)

df_letter_aux <- rbind(df_letter_aux, df_letter)

## ANOVA plots
ggplot(df, aes_string(x='basin_name', y=variable)) + 
  geom_boxplot() +
  geom_text(data=df_letter, aes(label = tukey_basin_name, x=basin_name, y=0.5),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot + 
  ylab(my_ylab) + xlab('')
```


```{r}
# Sen Pre
# ANOVA 
variable <- 'sen_slope_pre'
my_ylab <- 'Sen slope Pre'

myformula <- formula(paste(variable, '~ basin_name', sep=''))
mod <- aov(myformula, data=df)
pander(tidy(mod), caption=my_ylab)

## Multiple comparisons 
tuk <- glht(mod, linfct = mcp(basin_name = "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(basin_name = as.factor(lhs),
         tukey_basin_name = letters) %>%
  mutate(variable = variable)

df_letter_aux <- rbind(df_letter_aux, df_letter)

## ANOVA plots
ggplot(df, aes_string(x='basin_name', y=variable)) + 
  geom_boxplot() +
  geom_text(data=df_letter, aes(label = tukey_basin_name, x=basin_name, y=15),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot + 
  ylab(my_ylab) + xlab('')
```


# Pre Snow

```{r}
# Tau Pre Snow  
# ANOVA 
variable <- 'tau_pre_snow'
my_ylab <- 'Tau Pre Snow'

myformula <- formula(paste(variable, '~ basin_name', sep=''))
mod <- aov(myformula, data=df)
pander(tidy(mod), caption=my_ylab)

## Multiple comparisons 
tuk <- glht(mod, linfct = mcp(basin_name = "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(basin_name = as.factor(lhs),
         tukey_basin_name = letters) %>%
  mutate(variable = variable)

df_letter_aux <- rbind(df_letter_aux, df_letter)

## ANOVA plots
ggplot(df, aes_string(x='basin_name', y=variable)) + 
  geom_boxplot() +
  geom_text(data=df_letter, aes(label = tukey_basin_name, x=basin_name, y=0.7),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot + 
  ylab(my_ylab) + xlab('')
```


```{r}
# Sen Pre Snow 
# ANOVA 
variable <- 'sen_slope_pre_snow'
my_ylab <- 'Sen slope Pre Snow'

myformula <- formula(paste(variable, '~ basin_name', sep=''))
mod <- aov(myformula, data=df)
pander(tidy(mod), caption=my_ylab)

## Multiple comparisons 
tuk <- glht(mod, linfct = mcp(basin_name = "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(basin_name = as.factor(lhs),
         tukey_basin_name = letters) %>%
  mutate(variable = variable)

df_letter_aux <- rbind(df_letter_aux, df_letter)

## ANOVA plots
ggplot(df, aes_string(x='basin_name', y=variable)) + 
  geom_boxplot() +
  geom_text(data=df_letter, aes(label = tukey_basin_name, x=basin_name, y=20),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot + 
  ylab(my_ylab) + xlab('')
```

# Pre Snow per

```{r}
# Pre Snow per
# ANOVA 
variable <- 'tau_pre_snow_per'
my_ylab <- 'Tau Pre snow per'

myformula <- formula(paste(variable, '~ basin_name', sep=''))
mod <- aov(myformula, data=df)
pander(tidy(mod), caption=my_ylab)

## Multiple comparisons 
tuk <- glht(mod, linfct = mcp(basin_name = "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(basin_name = as.factor(lhs),
         tukey_basin_name = letters) %>%
  mutate(variable = variable)

df_letter_aux <- rbind(df_letter_aux, df_letter)

## ANOVA plots
ggplot(df, aes_string(x='basin_name', y=variable)) + 
  geom_boxplot() +
  geom_text(data=df_letter, aes(label = tukey_basin_name, x=basin_name, y=0.5),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot + 
  ylab(my_ylab) + xlab('')
```


```{r}
# Sen Pre snow per 
# ANOVA 
variable <- 'sen_slope_pre_snow_per'
my_ylab <- 'Sen slope Pre snow Per'

myformula <- formula(paste(variable, '~ basin_name', sep=''))
mod <- aov(myformula, data=df)
pander(tidy(mod), caption=my_ylab)

## Multiple comparisons 
tuk <- glht(mod, linfct = mcp(basin_name = "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(basin_name = as.factor(lhs),
         tukey_basin_name = letters) %>%
  mutate(variable = variable)

df_letter_aux <- rbind(df_letter_aux, df_letter)

## ANOVA plots
ggplot(df, aes_string(x='basin_name', y=variable)) + 
  geom_boxplot() +
  geom_text(data=df_letter, aes(label = tukey_basin_name, x=basin_name, y=2.5),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot + 
  ylab(my_ylab) + xlab('')
```

# Temp

```{r}
# Tau temp 
# ANOVA 
variable <- 'tau_temp'
my_ylab <- 'Tau Temp'

myformula <- formula(paste(variable, '~ basin_name', sep=''))
mod <- aov(myformula, data=df)
pander(tidy(mod), caption=my_ylab)

## Multiple comparisons 
tuk <- glht(mod, linfct = mcp(basin_name = "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(basin_name = as.factor(lhs),
         tukey_basin_name = letters) %>%
  mutate(variable = variable)

df_letter_aux <- rbind(df_letter_aux, df_letter)

## ANOVA plots
ggplot(df, aes_string(x='basin_name', y=variable)) + 
  geom_boxplot() +
  geom_text(data=df_letter, aes(label = tukey_basin_name, x=basin_name, y=0.7),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot + 
  ylab(my_ylab) + xlab('')
```


```{r}
# Sen temp 
# ANOVA 
variable <- 'sen_slope_temp'
my_ylab <- 'Sen slope Temp'

myformula <- formula(paste(variable, '~ basin_name', sep=''))
mod <- aov(myformula, data=df)
pander(tidy(mod), caption=my_ylab)

## Multiple comparisons 
tuk <- glht(mod, linfct = mcp(basin_name = "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(basin_name = as.factor(lhs),
         tukey_basin_name = letters) %>%
  mutate(variable = variable)

df_letter_aux <- rbind(df_letter_aux, df_letter)

## ANOVA plots
ggplot(df, aes_string(x='basin_name', y=variable)) + 
  geom_boxplot() +
  geom_text(data=df_letter, aes(label = tukey_basin_name, x=basin_name, y=.3),
           position = position_dodge(width=0.9)) +
  theme_bw() + mythemeggplot + 
  ylab(my_ylab) + xlab('')
```


```{r}
# Create a df with letter and variables
df_basin <- stats_basin %>%
  dplyr::inner_join(df_letter_aux, by=c('basin_name', 'variable'))

df_basin_tau <- df_basin %>%
  dplyr::filter(grepl("tau", variable))


ggplot(df_basin_tau, aes(x=basin_name, y=mean)) + 
  geom_point(stat='identity', size=3, fill='#222943') + 
  geom_errorbar(aes(ymin=mean - sd, ymax= mean + sd), 
                width=0.1,
                colour='#222943',
                position=position_dodge(.9)) +
  facet_wrap(~variable) + 
  geom_text(aes(label = tukey_basin_name, x=basin_name, y=mean),
           hjust=0, nudge_x = 0.2) + mythemeggplot +
  xlab('') + ylab('Taus')

```

```{r}
# Create a df with letter and variables
df_basin_sen <- df_basin %>%
  dplyr::filter(grepl("sen", variable))


ggplot(df_basin_sen, aes(x=basin_name, y=mean)) + 
  geom_point(stat='identity', size=3, fill='#222943') + 
  geom_errorbar(aes(ymin=mean - sd, ymax= mean + sd), 
                width=0.1,
                colour='#222943',
                position=position_dodge(.9)) +
  facet_wrap(~variable, scales="free") + 
  geom_text(aes(label = tukey_basin_name, x=basin_name, y=mean),
           hjust=0, nudge_x = 0.2) + mythemeggplot +
  xlab('') + ylab('Sen Slopes')
```


#### Tau Pre 
```{r}
pander(filter(df_basin_tau, variable=="tau_pre")) 
```

#### Sen Slope Pre
```{r}
pander(filter(df_basin_sen, variable=="sen_slope_pre")) 
```


#### Tau Pre Snow
```{r}
pander(filter(df_basin_tau, variable=="tau_pre_snow")) 
```

#### Sen Slope Pre Snow
```{r}
pander(filter(df_basin_sen, variable=="sen_slope_pre_snow")) 
```


#### Tau Pre Snow per
```{r}
pander(filter(df_basin_tau, variable=="tau_pre_snow_per")) 
```

#### Sen Slope Pre Snow per
```{r}
pander(filter(df_basin_sen, variable=="sen_slope_pre_snow_per")) 
```


#### Tau Temp
```{r}
pander(filter(df_basin_tau, variable=="tau_temp")) 
```

#### Sen Slope Temp
```{r}
pander(filter(df_basin_sen, variable=="sen_slope_temp")) 
```




# Explore trends by elevation 

## Taus 

```{r, fig.height=13 , fig.width=13}
myvariable <- 'tau_pre'
myylab <- 'Tau Pre'

# GAM 
myformula <- formula(paste(myvariable, '~ s(dem50mean)', sep=''))
modgam <- mgcv::gam(formula = myformula, 
                    data= df)
anova(modgam)
mygam_summ <- summary(modgam)
mygam_summ

# Plot GAM
a <-ggplot(df, aes_string(x='dem50mean', y=myvariable)) + geom_point(col='grey') + 
  geom_smooth(method="gam", formula = y ~ s(x), fill='red', col='red') + 
  annotate("text", x = 3000, y=0.3, 
           label= paste0("Dev. exp. = ", round((mygam_summ$dev.expl)*100, 2), " %")) + 
  mythemeggplot + 
  xlab('Elevation') + ylab(myylab)
  


myvariable <- 'tau_pre_snow'
myylab <- 'Tau Pre Snow'

# GAM 
myformula <- formula(paste(myvariable, '~ s(dem50mean)', sep=''))
modgam <- mgcv::gam(formula = myformula, 
                    data= df)
anova(modgam)
mygam_summ <- summary(modgam)
mygam_summ

# Plot GAM
b <- ggplot(df, aes_string(x='dem50mean', y=myvariable)) + geom_point(col='grey') + 
  geom_smooth(method="gam", formula = y ~ s(x), fill='red', col='red') + 
  annotate("text", x = 3000, y=0.6, 
           label= paste0("Dev. exp. = ", round((mygam_summ$dev.expl)*100, 2), " %")) + 
  mythemeggplot + 
  xlab('Elevation') + ylab(myylab)


myvariable <- 'tau_pre_snow_per'
myylab <- 'Tau Pre Snow per'

# GAM 
myformula <- formula(paste(myvariable, '~ s(dem50mean)', sep=''))
modgam <- mgcv::gam(formula = myformula, 
                    data= df)
anova(modgam)
mygam_summ <- summary(modgam)
mygam_summ

# Plot GAM
c <- ggplot(df, aes_string(x='dem50mean', y=myvariable)) + geom_point(col='grey') + 
  geom_smooth(method="gam", formula = y ~ s(x), fill='red', col='red') + 
  annotate("text", x = 3000, y=-0.25, 
           label= paste0("Dev. exp. = ", round((mygam_summ$dev.expl)*100, 2), " %")) + 
  mythemeggplot + 
  xlab('Elevation') + ylab(myylab)


myvariable <- 'tau_temp'
myylab <- 'Tau temp'

# GAM 
myformula <- formula(paste(myvariable, '~ s(dem50mean)', sep=''))
modgam <- mgcv::gam(formula = myformula, 
                    data= df)
anova(modgam)
mygam_summ <- summary(modgam)
mygam_summ

# Plot GAM
d <- ggplot(df, aes_string(x='dem50mean', y=myvariable)) + geom_point(col='grey') + 
  geom_smooth(method="gam", formula = y ~ s(x), fill='red', col='red') + 
  annotate("text", x = 3000, y=0.4, 
           label= paste0("Dev. exp. = ", round((mygam_summ$dev.expl)*100, 2), " %")) + 
  mythemeggplot + 
  xlab('Elevation') + ylab(myylab)


grid.arrange(a, d, b, c, nrow=2)
```

## Sen 

```{r, fig.height=13 , fig.width=13}
myvariable <- 'sen_slope_pre'
myylab <- 'Sen Slope pre'

# GAM 
myformula <- formula(paste(myvariable, '~ s(dem50mean)', sep=''))
modgam <- mgcv::gam(formula = myformula, 
                    data= df)
anova(modgam)
mygam_summ <- summary(modgam)
mygam_summ

# Plot GAM
a <-ggplot(df, aes_string(x='dem50mean', y=myvariable)) + geom_point(col='grey') + 
  geom_smooth(method="gam", formula = y ~ s(x), fill='red', col='red') + 
  annotate("text", x = 3000, y=10, 
           label= paste0("Dev. exp. = ", round((mygam_summ$dev.expl)*100, 2), " %")) + 
  mythemeggplot + 
  xlab('Elevation') + ylab(myylab)
  


myvariable <- 'sen_slope_pre_snow'
myylab <- 'Sen Slope Pre snow'

# GAM 
myformula <- formula(paste(myvariable, '~ s(dem50mean)', sep=''))
modgam <- mgcv::gam(formula = myformula, 
                    data= df)
anova(modgam)
mygam_summ <- summary(modgam)
mygam_summ

# Plot GAM
b <- ggplot(df, aes_string(x='dem50mean', y=myvariable)) + geom_point(col='grey') + 
  geom_smooth(method="gam", formula = y ~ s(x), fill='red', col='red') + 
  annotate("text", x = 3000, y=0, 
           label= paste0("Dev. exp. = ", round((mygam_summ$dev.expl)*100, 2), " %")) + 
  mythemeggplot + 
  xlab('Elevation') + ylab(myylab)


myvariable <- 'sen_slope_pre_snow_per'
myylab <- 'Sen slope Pre snow per'

# GAM 
myformula <- formula(paste(myvariable, '~ s(dem50mean)', sep=''))
modgam <- mgcv::gam(formula = myformula, 
                    data= df)
anova(modgam)
mygam_summ <- summary(modgam)
mygam_summ

# Plot GAM
c <- ggplot(df, aes_string(x='dem50mean', y=myvariable)) + geom_point(col='grey') + 
  geom_smooth(method="gam", formula = y ~ s(x), fill='red', col='red') + 
  annotate("text", x = 3000, y=2, 
           label= paste0("Dev. exp. = ", round((mygam_summ$dev.expl)*100, 2), " %")) + 
  mythemeggplot + 
  xlab('Elevation') + ylab(myylab)


myvariable <- 'sen_slope_temp'
myylab <- 'Sen Slope Temp'

# GAM 
myformula <- formula(paste(myvariable, '~ s(dem50mean)', sep=''))
modgam <- mgcv::gam(formula = myformula, 
                    data= df)
anova(modgam)
mygam_summ <- summary(modgam)
mygam_summ

# Plot GAM
d <- ggplot(df, aes_string(x='dem50mean', y=myvariable)) + geom_point(col='grey') + 
  geom_smooth(method="gam", formula = y ~ s(x), fill='red', col='red') + 
  annotate("text", x = 3000, y=0.1, 
           label= paste0("Dev. exp. = ", round((mygam_summ$dev.expl)*100, 2), " %")) + 
  mythemeggplot + 
  xlab('Elevation') + ylab(myylab)


grid.arrange(a, d, b, c, nrow=2)
```

