---
title: "Explore raw values of snow-cover indicators"
author: "AJ Perez-Luque (@ajpelu); FJ Bonet; J Herrero and R. Perez-Perez"
date: "2016 March"
output:  
    md_document:
      variant: markdown_github
---

```{r metadata, echo=FALSE}
# Set working directory 

machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/MS/CONGRESO_EGU2016/sn_snow_egu_2016', sep='')

```

```{r packages, warning=FALSE, message=FALSE}
# Load packages 
library("raster")
library("rgdal")
library("sp")
library("dplyr")
library("rasterVis") 
```

```{r, echo=FALSE}
# Set captions options
library("knitr")
opts_knit$set(eval.after='fig.cap')
cap <- function () {
    p <- last_plot()
    if (!is.null(p) && !is.null(p$labels$title)) {
        paste('Figure:', p$labels$title)
    } else {
        NULL # null caption
    }
}
```


## Prepare Data

* Read snow cover indicator data and subset snow cover duration 
* Read topographic data and position (spatial) data

```{r}
# Read data (snow cover)
snow <- read.csv(file=paste(di, "/data/raw/snow_sn.csv", sep= ""), header = TRUE) 
# --   

# Read spatial data and Get lat/long
centroides <- rgdal::readOGR(dsn=paste(di, "/data/geoinfo", sep=""),
                             layer = "centroides_selected", verbose = FALSE)
# Select only attributes of interest and rename them
centroides <- centroides[c("id")]

# Create lat/lng by id 
xycentroides <- cbind(centroides@data, coordinates(centroides))
names(xycentroides) <- c("nie_malla_modi_id", "lon","lat")

xycentroides <- filter(xycentroides, nie_malla_modi_id %in% snow$nie_malla_modi_id)

# -- 

# Read Topographic data 
rawtopo <- read.csv(file=paste(di, "/data/topo_nie_malla_modis.csv", sep=""),
                 header=TRUE,
                 sep = ",") 
# function to convert radian to degree 
rad2deg <- function(rad) {(rad * 180) / (pi)} 


topo <- rawtopo %>% 
  filter(id %in% snow$nie_malla_modi_id) %>% 
  mutate(nie_malla_modi_id = id, 
         slope50mean_deg = rad2deg(slope50mean),
         slope50median_deg = rad2deg(slope50median),
         aspect50mean_deg = rad2deg(aspect50mean),
         aspect50median_deg = rad2deg(aspect50median)) %>%
  dplyr::select(nie_malla_modi_id, dem50mean, dem50median, slope50mean_deg, 
         slope50median_deg, aspect50mean_deg, aspect50median_deg) 
# -- 


# Create objects with basis statistics for all indicators 
indicadores <- c("scd", "scod", "scmd", "scmc") 

# Loop to create objects with basis stats
for (i in indicadores) { 
  vnames <- c("nie_malla_modi_id", i) 
  aux <- snow %>%
    dplyr::select(one_of(vnames)) %>%
    mutate_(vinterest = i) %>%
    group_by_("nie_malla_modi_id") %>%
    summarise(mean=mean(vinterest),
              sd = sd(vinterest),
              cv = raster::cv(vinterest),
              se = sd / sqrt (length(vinterest))) %>%
    inner_join(topo, by="nie_malla_modi_id") %>%
    inner_join(xycentroides, by="nie_malla_modi_id") 
    
    assign(i, aux)
} 

```


### Spatial pattern of the snowcover indicators 

* Create raster maps of the summary stats for each indicator (`$indicator$`: `scd`, `scod`, `scmd`, `scmc`). Two raster maps will be created:

 * `r_mean_$indicator$`: mean values of the indicator for the pixel in the temporal serie.
 * `r_cv_$indicator$`: coefficient of variation of the indicator for the pixel in the temporal serie.

* Two additional raster maps will be created, with a mask of the elevation (those pixels above 1900 *m asl*). The names of the raster are: `r_mean_$indicator$_1900` and `r_cv_$indicator$_1900`. Pixels below 1900 masl show a value of `-1`. This value can be customized (change `updatevalue=-1` argument of the `mask` function). 

* All these rasters are stored at `./data/derived/` 

```{r} 
# Spatial data
# Reproject to utm and m
centroides <- spTransform(centroides, CRS("+init=epsg:23030"))

# Get projection 
projection(centroides) 

# Select only attributes of interest and rename them
centroides <- centroides[c("id")]
names(centroides) <-"nie_malla_modi_id"

# Which pixels are in the snow object
pix_comunes <- match(snow$nie_malla_modi_id, centroides$nie_malla_modi_id)

# Create spatial objetc with centroid of Sierra Nevada
centroides_sn <- centroides[pix_comunes,]


# Loop to create raster map 
for (i in indicadores) { 
  df <- get(i)
  
  # merge MKT and spatial pixel 
  aux_spatial <- sp::merge(x=centroides_sn, y=df, by="nie_malla_modi_id")
  aux_spatial_elev <- aux_spatial[c("nie_malla_modi_id", "dem50mean")]
  
  # raster auxiliar 
  aux_rast <- raster(aux_spatial, resolution=500)

  # raster of Mean values 
  mean_raster <- rasterize(aux_spatial, aux_rast, "mean", fun=mean)
  names(mean_raster) <- i # Set name of the raster layer 
  
  # raster of CV values
  cv_raster <- rasterize(aux_spatial, aux_rast, "cv", fun=mean)
  names(cv_raster) <- i 
  
  # raster of Elevation
  elev_raster <- rasterize(aux_spatial_elev, aux_rast, "dem50mean", fun=mean)
  elev_raster1900 <- elev_raster
  elev_raster1900[elev_raster1900 < 1900] <- NA
  
  # Mask by 1900
  mean_raster1900 <- mask(mean_raster, elev_raster1900, updatevalue=0)
  cv_raster1900 <- mask(cv_raster, elev_raster1900, updatevalue=0)
  
  
  # assign 
  name_mean <- paste("r_mean_",i, sep="")
  name_cv <- paste("r_cv_",i, sep="")
  name_mean1900 <- paste("r_mean_",i,"_1900", sep="")
  name_cv1900 <- paste("r_cv_",i,"_1900", sep="")
  
  assign(name_mean, mean_raster)
  writeRaster(mean_raster, file=paste(di, "/data/derived/r_mean_", i, ".asc", sep=""), overwrite=TRUE)
  
  assign(name_cv, cv_raster)
  writeRaster(cv_raster, file=paste(di, "/data/derived/r_cv_", i, ".asc", sep=""), overwrite=TRUE)
  
  assign(name_mean1900, mean_raster1900)
  writeRaster(mean_raster1900, file=paste(di, "/data/derived/r_mean_", i, "_1900", ".asc", sep=""), overwrite=TRUE)
  
  assign(name_cv1900, cv_raster1900)
  writeRaster(cv_raster1900, file=paste(di, "/data/derived/r_cv_", i, "_1900", ".asc", sep=""), overwrite=TRUE)
  }
```

## Visualization of the Snow Cover indicators 

### Snow Cover Duration

```{r, echo=FALSE, fig.cap='Figure 1', fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
# mypal <- brewer.pal(9, "Blues")
# Add a colour for 0 value
# mypal <- c("#D2D2D2", mypal)

mypal <- c("#D2D2D2", "#C6DBEF", "#9ECAE1", "#6BAED6", "#4292C6", "#2171B5", "#08519C", "#08306B")

# Specify the color palette
myTheme=rasterTheme(region=mypal)
# myTheme$panel.background$col <- 'gray'

# Mean plot
levelplot(r_mean_scd_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=c(0,seq(50,250,by=25)),
          main="Snow Cover Duration (mean values, days)")

```

```{r, echo=FALSE, fig.cap='Figure 1', fig.height=8 , fig.width=11}
# CV plot
mypal_cv <- brewer.pal(9, "Oranges")
mypal_cv <- c("#D2D2D2", mypal_cv)

myTheme_cv=rasterTheme(region=mypal_cv)

levelplot(r_cv_scd_1900, 
          par.settings=myTheme_cv, margin=list(axis=TRUE), colorkey=TRUE, 
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=seq(0,100,by=10),
          main="Snow Cover Duration (CV)")
```

