---
title: "Explore raw values of snow-cover indicators"
author: "AJ Perez-Luque (@ajpelu); FJ Bonet; J Herrero and R. Perez-Perez"
date: "2016 March"
output:
  md_document:
      variant: markdown_github
---

```{r metadata, echo=FALSE}
# Set working directory 

# machine <- 'ajpelu'
machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/MS/CONGRESO_EGU2016/sn_snow_egu_2016', sep='')

```

```{r packages, warning=FALSE, message=FALSE}
# Load packages 
library("raster")
library("rgdal")
library("sp")
library("dplyr")
library("rasterVis") 
library("multcomp")
library("broom") # tidy output models
library("grid") # multiple plots
library("gridExtra") # multiple plots
library("ggplot2")
```

```{r, echo=FALSE}
# AUTOCAPTIONS


# https://rpubs.com/ajlyons/autonumfigs
```


```{r auto_num_functions, echo=FALSE}
# A function for generating captions and cross-references

fig <- local({
  i <- 0
  list(
    cap=function(refName, text, center=FALSE, col="black", inline=FALSE) {
      i <<- i + 1
      ref[[refName]] <<- i
      css_ctr <- ""
      if (center) css_ctr <- "text-align:center; display:inline-block; width:100%;"
      cap_txt <- paste0("<span style=\"color:", col, "; ", css_ctr, "\">Figure ", i, ": ", text , "</span>")
      anchor <- paste0("<a name=\"", refName, "\"></a>")
      if (inline) {
        paste0(anchor, cap_txt)    
      } else {
        list(anchor=anchor, cap_txt=cap_txt)
      }
    },
    
    ref=function(refName, link=FALSE, checkRef=TRUE) {
      
      ## This function puts in a cross reference to a caption. You refer to the
      ## caption with the refName that was passed to fig$cap() (not the code chunk name).
      ## The cross reference can be hyperlinked.
      
      if (checkRef && !refName %in% names(ref)) stop(paste0("fig$ref() error: ", refName, " not found"))
      if (link) {
        paste0("<A HREF=\"#", refName, "\">Figure ", ref[[refName]], "</A>")
      } else {
        paste0("Figure ", ref[[refName]])
      }
    },
    
    ref_all=function(){
      ## For debugging
      ref
    })
})

```

```{r setup, echo=FALSE}
## This chunk replaces the default hook for processing plots. It achieves the purposes,
## of laying out auto-numbered captions, but other functionality may be gone.

library("knitr")
knit_hooks$set(plot = function(x, options) {
  sty <- ""
  if (options$fig.align == 'default') {
    sty <- ""
  } else {
    sty <- paste0(" style=\"text-align:", options$fig.align, ";\"")
  }
  
  if (is.list(options$fig.cap)) {
    ## options$fig.cap is a list returned by the function fig$cap()
    str_caption <- options$fig.cap$cap_txt
    str_anchr <- options$fig.cap$anchor
  } else {
    ## options$fig.cap is a character object (hard coded, no anchor)
    str_caption <- options$fig.cap
    str_anchr <- ""
  }
  
  paste('<figure', sty, '>', str_anchr, '<img src="',
        opts_knit$get('base.url'), paste(x, collapse = '.'),
        '"><figcaption>', str_caption, '</figcaption></figure>',
        sep = '')
  
})
```

```{r read_fig_cap, echo=FALSE}

## This chucnk will read through *this* Rmd file, and attempt to extract all of the 
## labels (not caption text) used for Figure captions. These labels are used
## as anchors, so scanning through the document now will allow us to create cross references
## before the caption actually appears. 

## Get the name of this Rmd file
rmdFn <- knitr::current_input()  # filename of input document

## Read lines and close connection
rmdCon <- file(rmdFn, open = "r")
rmdLines <- readLines(rmdCon)
close(rmdCon)

## Pull out all occurences of at least one back tick, followed 
## by any number of characters, followed by fig$cap (all on one line)
figscap_idx <- grep("`+(.*)fig\\$cap", rmdLines)
rmdLines <- rmdLines[figscap_idx]

## Get rid of everything up until the start of the caption label
## This presumes the caption label is the first argument of fig$cap()
## E.g., fig.cap = fig$cap("my_label", ...)
rmdLinesSansPre <- sub("(.*)fig\\$cap(.*?)[\"']", "", rmdLines)

## Identify everything up until the first quote
match_data <- regexpr("(.*?)[\"']", rmdLinesSansPre)

## Reduce the length by one, because we're not interested in the final quote
attr(match_data, "match.length") <- attr(match_data, "match.length") - 1

## Extract
fig_labels <- regmatches(rmdLinesSansPre, match_data, invert=FALSE)

if (length(fig_labels) > 0) {
  
  ## Test for duplicates
  if (anyDuplicated(fig_labels) > 0) stop("Duplicate caption labels detected")
  
  ## Create a named list of Figure numbers
  ref <- as.list(1:length(fig_labels))
  names(ref) <- fig_labels
}    
```


## Prepare Data

* Read snow cover indicator data and subset snow cover duration 
* Read topographic data and position (spatial) data

```{r}
# Read data (snow cover)
snow <- read.csv(file=paste(di, "/data/raw/snow_sn.csv", sep= ""), header = TRUE) 
# --   

# Read spatial data and Get lat/long
centroides <- rgdal::readOGR(dsn=paste(di, "/data/geoinfo", sep=""),
                             layer = "centroides_selected", verbose = FALSE)
# Select only attributes of interest and rename them
centroides <- centroides[c("id")]

# Create lat/lng by id 
xycentroides <- cbind(centroides@data, coordinates(centroides))
names(xycentroides) <- c("nie_malla_modi_id", "lon","lat")

xycentroides <- filter(xycentroides, nie_malla_modi_id %in% snow$nie_malla_modi_id)

# -- 

# Read Topographic data 
rawtopo <- read.csv(file=paste(di, "/data/topo_nie_malla_modis.csv", sep=""),
                    header=TRUE,
                    sep = ",") 
# function to convert radian to degree 
rad2deg <- function(rad) {(rad * 180) / (pi)} 


topo <- rawtopo %>% 
  filter(id %in% snow$nie_malla_modi_id) %>% 
  mutate(nie_malla_modi_id = id, 
         slope50mean_deg = rad2deg(slope50mean),
         slope50median_deg = rad2deg(slope50median),
         aspect50mean_deg = rad2deg(aspect50mean),
         aspect50median_deg = rad2deg(aspect50median)) %>%
  dplyr::select(nie_malla_modi_id, dem50mean, dem50median, slope50mean_deg, 
                slope50median_deg, aspect50mean_deg, aspect50median_deg) 
# -- 


# Create objects with basis statistics for all indicators 
indicadores <- c("scd", "scod", "scmd", "scmc") 

# Loop to create objects with basis stats
for (i in indicadores) { 
  vnames <- c("nie_malla_modi_id", i) 
  aux <- snow %>%
    dplyr::select(one_of(vnames)) %>%
    mutate_(vinterest = i) %>%
    group_by_("nie_malla_modi_id") %>%
    summarise(mean=mean(vinterest),
              sd = sd(vinterest),
              cv = raster::cv(vinterest),
              se = sd / sqrt (length(vinterest))) %>%
    inner_join(topo, by="nie_malla_modi_id") %>%
    inner_join(xycentroides, by="nie_malla_modi_id") 
  
  assign(i, aux)
} 

```


### Spatial pattern of the snowcover indicators 

* Create raster maps of the summary stats for each indicator (`$indicator$`: `scd`, `scod`, `scmd`, `scmc`). Two raster maps will be created:
  
  * `r_mean_$indicator$`: mean values of the indicator for the pixel in the temporal serie.
* `r_cv_$indicator$`: coefficient of variation of the indicator for the pixel in the temporal serie.

* Two additional raster maps will be created, with a mask of the elevation (those pixels above 1900 *m asl*). The names of the raster are: `r_mean_$indicator$_1900` and `r_cv_$indicator$_1900`. Pixels below 1900 masl show a value of `-1`. This value can be customized (change `updatevalue=-1` argument of the `mask` function). 

* All these rasters are stored at `./data/derived/` 

```{r} 
# Spatial data
# Reproject to utm and m
centroides <- spTransform(centroides, CRS("+init=epsg:23030"))

# Get projection 
projection(centroides) 

# Select only attributes of interest and rename them
centroides <- centroides[c("id")]
names(centroides) <-"nie_malla_modi_id"

# Which pixels are in the snow object
pix_comunes <- match(snow$nie_malla_modi_id, centroides$nie_malla_modi_id)

# Create spatial objetc with centroid of Sierra Nevada
centroides_sn <- centroides[pix_comunes,]


# Loop to create raster map 
for (i in indicadores) { 
  df <- get(i)
  
  # merge MKT and spatial pixel 
  aux_spatial <- sp::merge(x=centroides_sn, y=df, by="nie_malla_modi_id")
  aux_spatial_elev <- aux_spatial[c("nie_malla_modi_id", "dem50mean")]
  
  # raster auxiliar 
  aux_rast <- raster(aux_spatial, resolution=500)
  
  # raster of Mean values 
  mean_raster <- rasterize(aux_spatial, aux_rast, "mean", fun=mean)
  names(mean_raster) <- i # Set name of the raster layer 
  
  # raster of CV values
  cv_raster <- rasterize(aux_spatial, aux_rast, "cv", fun=mean)
  names(cv_raster) <- i 
  
  # raster of Elevation
  elev_raster <- rasterize(aux_spatial_elev, aux_rast, "dem50mean", fun=mean)
  elev_raster1900 <- elev_raster
  elev_raster1900[elev_raster1900 < 1900] <- NA
  
  # Mask by 1900
  mean_raster1900 <- mask(mean_raster, elev_raster1900, updatevalue=0)
  cv_raster1900 <- mask(cv_raster, elev_raster1900, updatevalue=0)
  
  
  # assign 
  name_mean <- paste("r_mean_",i, sep="")
  name_cv <- paste("r_cv_",i, sep="")
  name_mean1900 <- paste("r_mean_",i,"_1900", sep="")
  name_cv1900 <- paste("r_cv_",i,"_1900", sep="")
  
  assign(name_mean, mean_raster)
  writeRaster(mean_raster, file=paste(di, "/data/derived/r_mean_", i, ".asc", sep=""), overwrite=TRUE)
  
  assign(name_cv, cv_raster)
  writeRaster(cv_raster, file=paste(di, "/data/derived/r_cv_", i, ".asc", sep=""), overwrite=TRUE)
  
  assign(name_mean1900, mean_raster1900)
  writeRaster(mean_raster1900, file=paste(di, "/data/derived/r_mean_", i, "_1900", ".asc", sep=""), overwrite=TRUE)
  
  assign(name_cv1900, cv_raster1900)
  writeRaster(cv_raster1900, file=paste(di, "/data/derived/r_cv_", i, "_1900", ".asc", sep=""), overwrite=TRUE)
}
```

## Visualization of the Snow Cover indicators 

### Snow Cover Duration

```{r, echo=FALSE, fig.cap=fig$cap("scd_mean", "Mean values of Snow cover duration"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
# mypal <- brewer.pal(9, "Blues")
# Add a colour for 0 value
# mypal <- c("#D2D2D2", mypal)

mypal <- c("#D2D2D2", "#C6DBEF", "#9ECAE1", "#6BAED6", "#4292C6", "#2171B5", "#08519C", "#08306B")

# Specify the color palette
myTheme=rasterTheme(region=mypal)
# myTheme$panel.background$col <- 'gray'

# Mean plot
levelplot(r_mean_scd_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=c(0,seq(50,250,by=25)),
          main="Snow Cover Duration (mean values, days)")

```

```{r, echo=FALSE, fig.cap=fig$cap("scd_cv", "Coefficient of Variation of values of Snow cover duration"), fig.height=8 , fig.width=11}
# CV plot
mypal_cv <- brewer.pal(9, "Oranges")
mypal_cv <- c("#D2D2D2", mypal_cv)

myTheme_cv=rasterTheme(region=mypal_cv)

levelplot(r_cv_scd_1900, 
          par.settings=myTheme_cv, margin=list(axis=TRUE), colorkey=TRUE, 
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=seq(0,100,by=10),
          main="Snow Cover Duration (CV)")
```


### Snow Cover Onset Date

```{r, echo=FALSE, fig.cap=fig$cap("scod_mean", "Mean values of Snow cover Onset Date"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
mypal <- brewer.pal(9, "YlOrRd")
# Add a colour for 0 value
mypal <- c("#D2D2D2", mypal)
# Specify the color palette
myTheme=rasterTheme(region=mypal)


levelplot(r_mean_scod_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          #cut=20,
          at=c(0,seq(15,105,by=7)),
          main="Snow Cover Onset Date (mean values, hydrological day)") 

levelplot(r_mean_scod_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          main="Snow Cover Onset Date (mean values, hydrological day)") 
```

```{r, echo=FALSE, fig.cap=fig$cap("scod_cv", "Coefficient of Variation of values of Snow cover Onset date"), fig.height=8 , fig.width=11}
# CV plot
mypal_cv <- brewer.pal(9, "Oranges")
mypal_cv <- c("#D2D2D2", mypal_cv)

myTheme_cv=rasterTheme(region=mypal_cv)

levelplot(r_cv_scod_1900, 
          par.settings=myTheme_cv, margin=list(axis=TRUE), colorkey=TRUE, 
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=seq(0,100,by=10),
          main="Snow Cover Onset Date (CV)")
```

### Snow Cover Melting Date

```{r, echo=FALSE, fig.cap=fig$cap("scmd_mean", "Mean values of Snow cover Melting Date"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
mypal <- rev(brewer.pal(9, "Greens"))
# Add a colour for 0 value
mypal <- c("#D2D2D2", mypal)
# Specify the color palette
myTheme=rasterTheme(region=mypal)


levelplot(r_mean_scmd_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          #cut=20,
          at=c(0,seq(140,280,by=20)),
          main="Snow Cover Melting Date (mean values, hydrological day)") 

levelplot(r_mean_scmd_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          main="Snow Cover Melting Date (mean values, hydrological day)", 
          at=c(0,seq(140,280,by=5)))
```

```{r, echo=FALSE, fig.cap=fig$cap("scmd_cv", "Coefficient of Variation of values of Snow cover Melting date"), fig.height=8 , fig.width=11}
# CV plot
mypal_cv <- brewer.pal(9, "Oranges")
mypal_cv <- c("#D2D2D2", mypal_cv)

myTheme_cv=rasterTheme(region=mypal_cv)

levelplot(r_cv_scmd_1900, 
          par.settings=myTheme_cv, margin=list(axis=TRUE), colorkey=TRUE, 
          #contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          #at=seq(0,50,by=7.5),
          main="Snow Cover Melting Date (CV)")
```

### Snow Cover Melting Cycles

```{r, echo=FALSE, fig.cap=fig$cap("scmc_mean", "Mean values of Snow cover Melting Date"), fig.height=8 , fig.width=11}

# Select a palette
# http://colorbrewer2.org/
mypal <- brewer.pal(5, "Purples")
# Add a colour for 0 value
mypal <- c("#D2D2D2", mypal)
# Specify the color palette
myTheme=rasterTheme(region=mypal)


levelplot(r_mean_scmc_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          # Contour options
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=seq(0,5,by=1),
          main="Snow Cover Melting cycles (mean values, number)") 

levelplot(r_mean_scmc_1900, 
          par.settings=myTheme, margin=list(axis=TRUE), colorkey=TRUE, 
          main="Snow Cover Melting cycles (mean values, number)")
```

```{r, echo=FALSE, fig.cap=fig$cap("scmc_cv", "Coefficient of Variation of values of Snow cover Melting cycles"), fig.height=8 , fig.width=11}
# CV plot
mypal_cv <- brewer.pal(9, "Oranges")
mypal_cv <- c("#D2D2D2", mypal_cv)

myTheme_cv=rasterTheme(region=mypal_cv)

levelplot(r_cv_scmc_1900, 
          par.settings=myTheme_cv, margin=list(axis=TRUE), colorkey=TRUE, 
          contour=TRUE, labels=TRUE, pretty=TRUE, 
          # Breaks
          at=seq(0,100,by=10),
          main="Snow Cover Melting Date (CV)")
```

# Explore patterns of raw values

## Snow Cover Onset Dates 

* Analyze the effect of aspect, slope and longtitude on snow cover onset date

```{r, echo=FALSE}
# Subset data and classify elevation and aspect
scod1900 <- scod %>% filter(dem50mean > 1900) %>%
  mutate(dem50mean_group = cut(dem50mean, 
                         breaks = seq(from=0, to=3500, by=250),
                         labels = c("0-250", "251-500","501-750","751-1000",
                                    "1001-1250","1251-1500","1501-1750","1751-2000",
                                    "2001-2250","2251-2500","2501-2750", "2751-3000",
                                    "3001-3250", "3251-3500")),
         aspect50mean_deg_group = cut(aspect50mean_deg, 
                                      breaks= c(22.5, 67.5, 112.5, 157.5, 
                                                202.5, 247.5, 292.5, 337.5, 361),
                                      labels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW"))) 


### Model 
m_scod <- glm(mean ~ lon + aspect50mean_deg_group + slope50mean_deg, data=scod1900)
tidy(m_scod)
```

:red_circle: broom tidy glm factor level

```{r, echo=FALSE, fig.cap=fig$cap("scod_tukey_aspect", "Snow cover onset dates by aspect"), fig.height=8 , fig.width=11}
# Plot of Snow Cover Onset dates by aspect 

# Remove na level of aspect 
df <- scod1900 %>% na.omit() 

# Compute multiple comparisons 
tuk <- glht(m_scod, linfct = mcp(aspect50mean_deg_group= "Tukey"))
# Convert comparisons into letters 
df_letter <- fortify(cld(tuk)) %>%
  transmute(aspect50mean_deg_group = as.factor(lhs),
         tukey_aspect = letters) 

# Plot 
ggplot(df, aes(x=aspect50mean_deg_group, y=mean)) +geom_boxplot() + 
  geom_text(data=df_letter, aes(label = tukey_aspect, x=aspect50mean_deg_group, y=100),
           position = position_dodge(width=0.9)) +
  theme_bw() + 
  ylab('Snow cover onset date (hydrological day)') +
  xlab('aspect') + 
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.background=element_rect(fill='white')) 
```

```{r echo=FALSE, fig.cap=fig$cap("scod_tukey_aspect_effects", "Snow cover onset dates by aspect (effect sizes"), fig.height=8 , fig.width=11}
# Effects comparison (aspect)
CI <- confint(tuk)

ggplot(mapping = aes(lhs, estimate)) +
   geom_linerange(aes(ymin = lwr, ymax = upr), data = CI) +
   geom_point(aes(size = p), data = summary(tuk)) +
   scale_size(trans = "reverse") + theme_bw()
```


```{r echo=FALSE, fig.cap=fig$cap("scod_lon_slope", "Snow cover onset dates by Longitude and Slope"), fig.height=8 , fig.width=11}

a <- ggplot(scod1900, aes(x=lon, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method='lm') +
    theme_bw() + 
  ylab('Snow cover onset date (hydrological day)') +
  xlab('Longitude') + 
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.background=element_rect(fill='white')) 
  

b <- ggplot(scod1900, aes(x=slope50mean_deg, y= mean)) + 
  geom_point(col='gray') + 
  geom_smooth(method='lm') +
    theme_bw() + 
  ylab('Snow cover onset date (hydrological day)') +
  xlab('Slope (degrees') + 
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.background=element_rect(fill='white')) 

grid.arrange(a, b, ncol=2)
```
















  
  
```{r, echo=FALSE, eval=FALSE}
ggplot(df, aes(x=aspect, y=mean)) + 
  geom_bar(position=position_dodge(), stat="identity", fill='#084594') +
  geom_errorbar(aes(ymax = mean + sd, ymin=mean - sd),
                  width=.25, col= '#084594', 
                  position=position_dodge(.9)) + 
  theme_bw() + 
  ylab('Snow cover onset date (hydrological day)') +
  xlab('aspect') + 
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.background=element_rect(fill='white')) 
```





