---
title: "Compute Mann-Kendall Theil Sen of snow cover"
author: "AJ Perez-Luque (@ajpelu); FJ Bonet; J Herrero and R. Perez-Perez"
date: "2016 March"
output:  
    md_document:
      variant: markdown_github
---

```{r metadata, echo=FALSE}
################################################################
# Set working directory 

machine <- 'ajpelu'
# machine <- 'ajpeluLap'
di <- paste('/Users/', machine, '/Dropbox/MS/CONGRESO_EGU2016/sn_snow_egu_2016', sep='')
################################################################
```

En primer lugar leemos los datos
```{r packages, warning=FALSE}
################################################################
# Load packages 
library("dplyr")
library("modeest")
################################################################

################################################################
# Read data

# Define name of indicators (see variables names)
indicadores <- c("scd", "scod", "scmd", "scmc", "pre", "pre_snow", "pre_snow_per", "temp")

# Loop to read files 
for (j in indicadores){ 
  aux <- read.csv(file=paste(di, "/data/derived/", j, ".csv", sep= ""),
              header = TRUE,
              sep = ',')
  assign(j, aux)
}


# Read Topographic data 
rawtopo <- read.csv(file=paste(di, "/data/topo_ni_malla_modi_id.csv", sep=""),
                 header=TRUE,
                 sep = ",") 


topoaux <- rawtopo %>% 
  select(nie_malla_modi_id, altura, twi, pendiente) 

topo <- topoaux %>% 
  group_by(nie_malla_modi_id) %>% 
  summarise_each(funs(mode=mlv(., method='mfv')[['M']],
                      mean), altura:pendiente)



test <- topoaux[topoaux$nie_malla_modi_id == 1752781,]

mean(test$altura)
mlv(test$altura, method = 'mfv')





%>%
  group_by(nie_malla_modi_id) %>% 
  mutate_each_(funs(media= mean(.)), altura:pendiente)


%>%
  group_by(nie_malla_modi_id) %>%
  mutate_each(funs(media = mean(.)), c(altura, twi, pendiente))




              
xx <- c(1,2,2,2,3,4,2,3,1)
mfv(xx)
mlv()



# Join Wimmed data
join_aux1 <- inner_join(PnXpxXyear, PorcPnXpxXyear, by=c('id','year'))
join_aux2 <- inner_join(PXpxXyear, TXpxXyear, by=c('id','year'))
join_aux3 <- inner_join(join_aux1, join_aux2, by=c('id','year'))

# Set rigth names
names(join_aux3)[1] <- "nie_malla_modi_id"
names(snow_sn)[3] <- "year"

# Join with snow data
mydf <- inner_join(snow_sn, join_aux3, by=c("nie_malla_modi_id", "year"))

# remove temp files
rm(join_aux1, join_aux2, join_aux3)
```

A continuación creamos un bucle para obtener de cada indicador la tendencia y la pendiente. Aplicamos la técnica de Mann-Kendall-Theil-Sen. El bucle realiza las siguientes operaciones: 

* Crea una variable con los indicadores para el bucle. 
* Realiza un subset de datos con los valores del indicador para cada pixel. Obtenemos una serie de datos por cada pixel. 
* Calculamos la tendencia y la pendiente
* Guardamos los resultamos como csv (uno por cada indicador, ver `./data/derived/`)

```{r}
# Loop to compute the MKT by indicator

# Define name of indicators (see variables names)
indicadores <- c("scd", "scod", "scmd", "scmc", "pre", "pre_snow", "pre_snow_per", "temp")

# subset by indicator
for (j in indicadores){ 
  # Create a subset by indicator
  subdf <- mydf[, names(mydf) %in% c("nie_malla_modi_id", "year", j )]
  
  # Manipule data. Transpose 
  # la funcion aggregate es porque tenemos duplicados 
  aux <- dcast(subdf, year ~ nie_malla_modi_id, value.var = j, fun.aggregate = mean) 

  # Create a zoo object 
  auxts <- zoo(aux[-1],aux[,1])
  
  # Compute Mann Kendall and Theil 
  theil <- mannKen(as.ts(auxts))
  
  # Prepare data output
  theil <- as.data.frame(theil)
  
  theil_df <- theil %>% 
    mutate(nie_malla_modi_id = as.numeric(rownames(theil))) %>%
    mutate(sen_slope = round(sen.slope, 3),
           sen_slope_per = round(sen.slope.pct, 2),
           p_value = round(p.value, 5),
           tau = round(tau, 3)) %>%
    dplyr::select(nie_malla_modi_id, sen_slope, sen_slope_per, p_value, tau)
  
  # As data.frame
  theil_df <- as.data.frame(theil_df)
  
  # Rename variables adding dataframe name
  
  # Create vector of new names 
  new_names <- c(paste(names(theil_df), j ,sep='_'))
  
  theil_rename <- theil_df %>% rename_(.dots = setNames(names(.), new_names)) %>%
    dplyr::select(nie_malla_modi_id = starts_with("nie_malla_modi_id"), everything())
  
  # Write table 
  write.table(assign(j, theil_rename), file=paste(di, "/data/derived/", j, ".csv", sep=""), row.names=FALSE, sep=',')
  
  # Check Results. Get the pixel with NA in pvalue. Save results
  aux_na <- theil_df[!complete.cases(theil_df),]
  aux_na$var <- rep(j,nrow(aux_na))
  write.table(aux_na[,c("nie_malla_modi_id","var")], file=paste(di, "/data/derived/na_", j, ".csv", sep=""), row.names=FALSE, sep=',')
} 
```





```{r}
#### Read topo data 

# Create a dataframe with all variables and add Elevation data
### Add elevation data
library("rgdal")
library("sp")


# The pixles polygon shape is in 4326 
# Read pixels vectorial layer 
pixels <- rgdal::readOGR(dsn = paste(di, "/data/geoinfo/", sep=""), 
                         layer = "nie_malla_modis_selected", 
                         verbose = FALSE)

pixels <- pixels[, c("id","elevacion")]
names(pixels) <- c("nie_malla_modi_id","elevacion")

## Join data of Elevation with mk data
elev <- as.data.frame(pixels@data)

## Create interval variables (250 m)
elev <- elev %>% 
  mutate(elevGroup = cut(elevacion, 
                         breaks = seq(from=0, to=3500, by=250),
                         labels = c("0-250", "251-500","501-750","751-1000",
                                    "1001-1250","1251-1500","1501-1750","1751-2000",
                                    "2001-2250","2251-2500","2501-2750", "2751-3000",
                                    "3001-3250", "3251-3500"))) 

# Create un dataframe con todos los datos
fulldf <- elev %>% 
  inner_join(scod,  by=c("nie_malla_modi_id")) %>% 
  inner_join(scd,  by=c("nie_malla_modi_id")) %>% 
  inner_join(scmd,  by=c("nie_malla_modi_id")) %>% 
  inner_join(scmc,  by=c("nie_malla_modi_id")) %>% 
  inner_join(pre,  by=c("nie_malla_modi_id")) %>% 
  inner_join(pre_snow,  by=c("nie_malla_modi_id")) %>% 
  inner_join(pre_snow_per,  by=c("nie_malla_modi_id")) %>% 
  inner_join(temp,  by=c("nie_malla_modi_id")) 
  

### Create subset of data > 1900 elev and only taus

taus <- fulldf %>%
  dplyr::select(nie_malla_modi_id, 
                elevacion,
                starts_with("tau"))


tau1900 <- fulldf %>% 
  filter(elevacion > 1900) %>% 
  dplyr::select(nie_malla_modi_id, elevacion, elevGroup, starts_with("tau"))

ss1900 <- fulldf %>% 
  filter(elevacion > 1900) %>% 
  dplyr::select(nie_malla_modi_id, elevacion, elevGroup, starts_with("sen_slope"))

tau <- fulldf %>% 
  dplyr::select(nie_malla_modi_id, elevacion, elevGroup, starts_with("tau"))

```


```{r}
library(ggplot2)




plot(tau1900)



ggplot(tau, aes(tau_temp, tau_scmd)) + geom_point() + 
  facet_wrap(~elevGroup)

ggplot(tau1900, aes(elevacion,tau_scd)) + geom_point()
  geom_point(aes(colour=elevGroup))
  
ggplot(ss1900, aes(elevacion,sen_slope_pre_snow_per)) + geom_point()


```{r}
# Merge dataframe of interest with spatial information 
mm <- sp::merge(pixels, taus, by="nie_malla_modi_id")

# Select only the pixels of Sierra Nevada
mm1 <- mm[mm$nie_malla_modi_id %in% taus$nie_malla_modi_id, ]


r <- raster()
tt <- rasterize(mm1, field= 'tau_scmd')
projection(tt)
extent(tt)s <- extent(mm1)


tt <-rasterize(mm1)


# Rasterizar con coordenadas utm
raster_tendencia<-rasterFromXYZ(malla_puntos_final_aux)
plot(raster_tendencia)












# Loop to add elevation data and to create subsets with elevation > 1900 
for (k in indicadores){
  auxmk <- get(k) # Get object
  aux_elev <- inner_join(elev, auxmk, by="nie_malla_modi_id") # Join elevation data
  
  aux_elev1900 <- aux_elev %>% filter(elevacion > 1900)
  
  assign(k, aux_elev)
  assign(paste(k,'1900',sep=''), aux_elev1900)
  }















  
  












# Asignamos proyección al raster anterior
crs(raster_tendencia)<-"+proj=utm +zone=30"

# reproyectar a geográficas
raster_tendencia<-projectRaster(raster_tendencia, crs = "+proj=longlat")

# creamos máscara para recortar y remuestrear. Le ponemos la resolución y la extensión de raster_tendencia
mascara<-raster(crs = "+proj=longlat")
extent(mascara)<-extent(raster_tendencia)
res(mascara)<-c(0.00418,0.00418)

raster_tendencia<-resample(raster_tendencia, mascara)
plot(raster_tendencia)




plot(mm1)



head(pixels)
summary(pixels)

extent(pixels)
projection(pixels)


head(pixels)





